\documentclass[11pt]{amsart}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[colorlinks=true, pdfstartview=FitV, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}

\usepackage[centering]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{tikz,caption}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\usepackage{enumitem}
\setlist[itemize]{leftmargin=2em}
\setlist[enumerate]{leftmargin=2em}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage[aligntableaux=center, smalltableaux]{ytableau}
\usepackage{youngtab}
\usepackage{verbatim}
\usepackage{faktor}
\usepackage{xspace}
\usepackage{esvect}
\usepackage{stmaryrd}

\newtheorem{thm}{Theorem}

\definecolor{darkblue}{rgb}{0.0,0,0.7} % darkblue color
\definecolor{darkred}{rgb}{0.7,0,0} % darkred color
\definecolor{darkgreen}{rgb}{0, .6, 0} % darkgreen color

% Dark red emphasis
\newcommand{\defncolor}{\color{darkred}}
\newcommand{\defn}[1]{{\defncolor\emph{#1}}} % emphasis of a definition
\newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{\texttt{arXiv:#1}}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{property}[theorem]{Property}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{question}[theorem]{Question}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{problem}[theorem]{Problem}
\numberwithin{equation}{section}

\def\NN{{\mathbb N}}
\def\CC{{\mathbb C}}
\def\ZZ{{\mathbb Z}}

\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\st}{st}

\newcommand{\FGCCHA}{\textsf{FGCCHA}\xspace}
\newcommand{\FGCCHAs}{\textsf{FGCCHA}s\xspace}
\newcommand{\vecdim}{\overrightarrow{\dim}}
\newcommand{\veccard}{\overrightarrow{\operatorname{card}}}
\newcommand{\OPG}{\operatorname{OPG}}

\usepackage[colorinlistoftodos]{todonotes}

\newcommand{\mike}[1]{\todo[size=\tiny,color=green!30]{#1 \\ \hfill --- Mike}}
\newcommand{\lucas}[1]{\todo[size=\tiny,color=red!50]{#1 \\ \hfill --- Lucas}}
\newcommand{\felix}[1]{\todo[size=\tiny,color=Cyan]{#1 \\ \hfill --- Félix}}
\newcommand{\farhad}[1]{\todo[size=\tiny,color=blue!50]{#1 \\ \hfill --- Farhad}}
\newcommand{\eric}[1]{\todo[size=\tiny,color=BurntOrange!50]{#1 \\ \hfill --- Eric}}
\newcommand{\nicolas}[1]{\todo[size=\tiny,color=purple!50]{#1 \\ \hfill --- Nicolas}}
\setlength{\marginparwidth}{28mm}

\title{Free, graded, connected, co-commutative Hopf algebras are integer sequences}
\author{}

\begin{document}
\maketitle

\section{Introduction}
\lucas{This is an outline-by-demonstration: a sketch of what points to hit in order to tell this story right.  Not  an attempt at the final version.}
This paper builds a comprehensive dictionary between integer sequences and a class of Hopf algebras that we will call \FGCCHAs: \textsf{F}ree, \textsf{G}raded, \textsf{C}onnected, \textsf{C}ocommutative \textsf{H}opf \textsf{A}lgebras.  
A recent result of Aliniaeifard and Thiem~\cite{AT20} shows that the isomorphism class of an \FGCCHA $H$ can be fully determined by the \emph{dimension sequence}
\[
\vecdim(H) = \big( \dim(H_{1}), \dim(H_{2}), \dim(H_{3}), \ldots \big)
\qquad\text{where}\qquad
H = \bigoplus_{n \ge 0} H_{n}.
\]
This inspired us to ask: ``how much about \FGCCHAs can one say using only integer sequences?'' and we found the answer to be ``almost everything.''

\mike{example here?  Connes-Kreimer Hopf algebra of rooted trees is an \FGCCHA by Theorem 3 of \cite{CK99}
Theorem 3 states it is the dual of the enveloping algebra of the Lie algebra $\mathcal{L}_1$. To discuss: is it clear
from that presentation that it is free?}

Paragraph here about: why \FGCCHAs, why asking about isomorphism, maps between them make sense.  
Summarize NCSym story in 1-2 sentences (since this is one of the bigger \FGCCHAs) including Advances Paper.  Touch on other examples of \FGCCHAs and maps in combinatorics: (Bergeron Adv. Appl. paper, the tree paper, maybe look for some other examples).  Takeaway: these are nontrivial and interesting objects.

We find that essentially all categorical data about \FGCCHAs is determined by the dimension sequence via certain simple arithmetic transformations.  
This is reminiscent of Foissy's work on free and cofree Hopf algebras~\cite{F12}, but we are still
surprised by the definitive nature of our results on \FGCCHAs and expect them to be useful tools in future work.
\begin{enumerate}[itemsep = 1ex]
\item We give an elementary description of all isomorphism classes of \FGCCHA:
Theorem~\ref{thm:a sequence} states that an arbitrary nonnegative sequence $\vec{h}$
is the dimension sequence for an isomorphism class of \FGCCHAs if and only if it
remains nonnegative under a simple arithmetic transformation $\phi_{h, a}$
that we define in Section~\ref{sec:SequencePrelims}.

\item W show that every sub-Hopf algebra of an \FGCCHA is itself an \FGCCHA
and classify these using a second transformation of the dimension sequence
that we  denote by $\phi_{h, p}$.  Theorem~\ref{thm:subclassification2} states
that there is an injective homomorphisms $H \to K$ between \FGCCHAs $H$ and $K$
if and only if $\phi_{h, p}(\vecdim(H))$ is dominated by $\phi_{h, p}(\vecdim(K))$
term-wise, while Theorem~\ref{} give an algorithm to construct all such maps.

\item While not every quotient of an \FGCCHA is free, we determine when one \FGCCHA is the quotient of another.  With the transformation $\phi_{h, a}$ from (1), Theorem~\ref{} shows that there is a surjective homomorphisms $H \to K$ between \FGCCHAs $H$ and $K$ if and only if $\phi_{h, p}(\vecdim(H))$ dominates $\phi_{h, p}(\vecdim(K))$ term-wise, and Theorem~\ref{} gives an algorithm to produce all such maps.  \lucas{This is what I was talking about in Google Chat recently.  Keep?}

\item Any map between two \FGCCHAs factors naturally as the  composition of a surjection and injections.  Section~\ref{} uses (2) and (3) to classify all maps between \FGCCHAs.

\end{enumerate}
These results follow primarily from new applications of known results, though in a few cases there are gaps or ``folklore'' results with no written proof and we introduce truly novel proof methods.  
Throughout, we emphasize clear, constrictive statements which can be applied ``out of the box'' in the ongoing study of \FGCCHAs, with numerous examples.

We close with a discussion of our initial motivation on this project, demonstrating one application of our results.  
The \FGCCHA of symmetric functions in non-commuting variables $\mathsf{NCSym}$ has a number of natural Hopf subalgebras whose dimension sequence was equal to the Catalan sequence~\cite[A000108]{OEIS}; see Example~\ref{}.  
We sought to understand ``how many'' of these Catalan sub-Hopf algebras there were.  

\begin{problem}
\label{prob1}
Identify all of the Hopf subalgebras $C \subseteq \mathsf{NCSym}$ for which $\vecdim(C)$ is the sequence of Catalan numbers.
\end{problem}

With the information above, however, this is easy to resolve: (2) tells us that every one of these Catalan sub-Hopf algebras is an \FGCCHA, so Aliniaeifard-Theim~\cite{AT20} tell us they are isomorphic.  Moreover, Theorem~\ref{} gives an algorithm for constructing every such sub-Hopf algebra from some preliminary data about $\mathsf{NCSym}$.  This led us to the following, more ambitious question.

\begin{problem}
\label{prob2}
Classify all (isomorphism classes of) sub-Hopf algebras of $\mathsf{NCSym}$.
\end{problem}

Here, (1) comes into play: while (2) tells us how to check is a given \FGCCHA appears as a sub-Hopf algebra, we also need to be able to exhaustively enumerate \FGCCHAs to answer this question.  
We resolve a more general version of this question in Theorem~\ref{}, and Problem~\ref{prob2} specifically in Example~\ref{}.  

%At this point we turned to two 
%
%At this point we 
%
%We also solved this and answered it in terms of sequences.  In the process, we encountered two more fundamental questions about  \FGCCHAs.
%
%\begin{problem}
%Identify conditions on the graded dimensions that imply the existence of a free graded connected co-commutative Hopf algebra.
%\end{problem}
%
%\begin{problem}
%Given two Hopf algebras that
%have the same graded dimension, find a procedure for
%determining an (all) isomorphism(s) between the Hopf
%algebras.
%\end{problem}
%
%We state the answer to this problem
%as a calculation that one can perform on the
%sequence of graded dimensions.
%The answer (stated in Theorem \ref{}) is that one should calculate a
%related sequence representing the number of free generators
%of the Hopf algebra.
%If the sequence is non-negative, then we can
%construct a Hopf algebra with these graded dimensions.
%This provides a complete characterization of Hopf algebras
%of this type.  
%Every sequence of generators determines completely the Hopf algebra.

\section{Notation}

A \defn{composition} of $n$ is a sequence $\beta = (\beta_1, \beta_2, \ldots, \beta_\ell)$ of positive integers such that $\beta_{1} + \beta_{2} + \cdots + \beta_{\ell} = n$.  
We refer the the integers $\beta_{i}$ as the parts of $\beta$ and write $\ell(\beta)$ for the length of $\beta$, which is the number of parts.  
We will use the notation $\beta \vDash n$ to indicate that $\beta$ is a composition of $n$.

We say that $\lambda$ is a \defn{partition} of $n$, if $\lambda \vDash n$ and
$\lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_{\ell(\lambda)}$.  We will indicate that $\lambda$ is a partition
of $n$ with the notation $\lambda \vdash n$.  We will also use the notation $m_d(\lambda)$ to be the number of times
that $d$ appears as a part in $\lambda$.

Let $\CC^{\NN}$ denote the space of infinite sequences $(c_{n})_{n \ge 1} = (c_{1}, c_{2}, \ldots)$ of complex numbers, i.e.~$c_{n} \in \CC$ for all $n \ge 1$.  
Given two sequences $\vec{c} = (c_{1}, c_{2}, \ldots)$ and $\vec{d} = (d_{1}, d_{2}, \ldots)$, we write
\[
\vec{c} \le \vec{d} 
\qquad\text{if and only if}\qquad
\text{$c_{n} \le d_{n}$ for all $n \ge 1$}.
\]
Let $\vec{0} \in \CC^{\NN}$ denote the zero sequence, so that $\vec{c} \ge \vec{0}$ is and only if $\vec{c}$ consists of entirely nonnegative entries.  
For a composition $\alpha = (\alpha_{1}, \alpha_{2}, \ldots, \alpha_{\ell}) \vDash n$, define 
\[
c_{\alpha} = c_{\alpha_{1}} c_{\alpha_{2}} \cdots c_{\alpha_{\ell}}.
\]
Note that $c_{\alpha} = c_{\beta}$ whenever $\beta \vDash n$ is a composition with the same parts as $\alpha$ in a possibly different order, i.e.~ $\beta_{i} = \alpha_{\sigma(i)}$ for some permutation $\sigma$ of the integers $\{1,2, \ldots, \ell(\alpha)\}$.


%Given a sequence $\vec{a}$ consisting of nonnegative integers, let 
%\[
%X(\vec{a}) = \biguplus_{n \ge 1} \{x^{(n)}_{i} \;|\; 1 \le i \le a_{n} \}.
%\]
%and let $\mathfrak{L}(\vec{a})$ denote the free graded Lie algebra on the graded set $X(\vec{a})$.  
%
%For a graded Lie algebra $L$, $\mathcal{U}(L)$ will denote the universal enveloping algebra of $L$. We can define $\mathcal{U}(L)$ as a quotient of the tensor algebra $T(L)/I$ where $I$ is the ideal generated by $x \otimes y - y \otimes x - [x,y]$ for $x,y \in L$. Because $I$ is generated by homogeneous elements, $\mathcal{U}(L)$ inherits a grading from $L$. The Poincar\'{e}-Birkhoff-Witt Theorem implies that for a given ordered basis $\{x_i\}$ of $L$, a basis for $\mathcal{U}(L)$ is given by $\{x_{i_1}x_{i_2}\cdots x_{i_k}:i_1 \leq i_2 \leq \cdots \leq i_k\}.$

%For a graded Hopf algebra $H$, $\mathcal{P}(H)$ will denote graded Lie algebra of primitives, so that
%\[
%\mathcal{P}(H) = \{x \in H \;|\; \Delta(x) = 1 \otimes x + x \otimes 1\}.
%\]
%\lucas{Add reference/definition for universal enveloping algebra and primitives.  Add statement of Milnor--Moore and PBW theorem.}

For a graded vector space $V = \bigoplus_{n \ge 0} V_{n}$ with $V_{n} = \CC$, let
\[
\vecdim(V) = \left( \dim(V_{n}) \right)_{n \ge 1}.
\]
Similarly, for a graded set $X = \biguplus_{n \ge 1} X^{(n)}$, let
\[
\veccard(X) = \left( |X^{(n)}| \right)_{n \ge 0}
\]

\section{Combinatorics of sequences of numbers}
\label{sec:SequencePrelims}

Our main results make use of three interrelated sequences that we will denote by $\vec{h}$, $\vec{a}$, and $\vec{p}$.  
In this section we take a purely enumerative perspective to these sequences, assuming only that they satisfy the relation given in Equation~\eqref{eq:gf_relation} and may have values from a field of characteristic $0$.

However, in later sections these sequences come from an \FGCCHA $H$ and have an interpretation when they are non-negative integers.
This interpretation will provide useful motivation here:
\begin{itemize}
\item $\vec{h} = (h_{1}, h_{2}, \ldots)$ will be the graded dimensions of $H$,

\item $\vec{a} = (a_{1}, a_{2}, \ldots)$ will be the graded numbers of free generators of the algebra, and 

\item $\vec{p} = (p_{1}, p_{2}, \ldots)$ will be the graded dimension of the Lie algebra of primitives $\mathcal{P}(H)$.

\end{itemize}

To define our sequences, we make use of the fact---recorded in Proposition~\ref{prop:sequences} below---that any formal power series in $\CC[[t]]$ with constant term $1$ can be expressed in three equivalent ways,
\begin{equation}
\label{eq:gf_relation}
1 + \sum_{k \geq 1} h_k t^k = \frac{1}{1 - \sum_{m \geq 1} a_m t^m} = \prod_{d \geq 1} \frac{1}{(1-t^d)^{p_d}}~.
\end{equation}
which determines a triple of sequences: $(\vec{h}, \vec{a}, \vec{p})$ with $\vec{h} = (h_{1}, h_{2}, \ldots)$, $\vec{a} = (a_{1}, a_{2}, \ldots)$, and $\vec{p} = (p_{1}, p_{2}, \ldots)$.  

\begin{example}
Take $f(t) = 1 + 2t + 3 t^{2} + \cdots \in \CC[[t]]$, so that $\vec{h} = (2, 3, 4, \ldots)$.  Then $f(t)$ is a geometric series,
\[
f(t) = \frac{1}{1 - 2 t^{2} + t^{2}} = \frac{1}{(1-t)^{2}}
\]
so the remaining sequences are $\vec{a} = (2, -1, 0, \ldots)$ and $\vec{p} = (2, 0, 0, \ldots)$.
\end{example}

We will make extensive use of explicit formulas relating the sequences $\vec{h}$, $\vec{p}$, and $\vec{a}$.  While these are well-known, we include a proof for the sake of completeness.

\begin{proposition}
\label{prop:sequences}
Any one sequence $\vec{h}$, $\vec{a}$, or $\vec{p} \in \CC^{\NN}$ belongs to a unique triple $(\vec{h}, \vec{a}, \vec{p})$ of sequences that satisfy Equation~\eqref{eq:gf_relation}, given by:
\begin{enumerate}[label = (\roman*), itemsep = 1em]
\item $\displaystyle h_{n}
= \sum_{\beta \vDash n} a_\beta
= \sum_{\lambda \vdash n} \prod_{d \geq 1} \binom{p_d + m_d(\lambda) -1}{m_d(\lambda)}$, 

\item $\displaystyle a_n
= \sum_{\beta \vDash n} (-1)^{\ell(\beta)-1} h_\beta
= \sum_{\lambda \vdash n} (-1)^{\ell(\lambda)-1} \prod_{d \geq 1} \binom{p_d}{m_d(\lambda)}$, and

\item $\displaystyle p_n
= \sum_{d|n} \sum_{\beta \vDash d} \frac{d\cdot \mu(n/d)}{n \cdot \ell(\beta)} a_\beta
= \sum_{d|n} \sum_{\beta \vDash d} \frac{d\cdot \mu(n/d) (-1)^{\ell(\beta)-1}}{n \cdot \ell(\beta)} h_\beta$.

\end{enumerate}
\end{proposition}

\begin{proof}
%We will make use of the identities
%\[
%(1-t)^p = \sum_{k \geq 0} \binom{p}{k} t^k, \quad
%\frac{1}{(1-t)^p} = \sum_{k \geq 0} \binom{p+k-1}{p-1} t^k, \quad\text{and}\quad
%\log(1-t) = \sum_{j \geq 1} \frac{t^j}{j}
%\]
%to derive equations (i)--(iii) from Equation \eqref{eq:gf_relation}.

In order to see (i), take the geometric series expansion of the second and third expressions in Equation \eqref{eq:gf_relation} to obtain
\begin{equation}\label{eq:gf_relation_reciprocal}
1 + \sum_{k \geq 1} h_k t^k
= \sum_{k \geq 0} \left(\sum_{m \geq 1} a_m t^m\right)^k
= \prod_{d \geq 1} \left(\sum_{k \geq 0} \binom{p_d+k-1}{k} t^{kd}\right).
\end{equation}
Expanding each product of sums and isolating the coefficient of $t^{n}$ in each expression gives the desired equation.

Now we prove (ii).  Take the reciprocal of Equation \eqref{eq:gf_relation}
and isolate the sum over $m$ to obtain
\[
1 - \frac{1}{1 + \sum_{k \geq 1} h_k t^k} 
= \sum_{m \geq 1} a_m t^m 
= 1 - \prod_{d \geq 1} (1-t^d)^{p_d}.
\]
Expanding the left- and rightmost expressions yields
\[
\sum_{r \geq 1} (-1)^{r-1} \left(\sum_{k \geq 1} h_k t^k\right)^r
= \sum_{m \geq 1} a_m t^m
= 1 - \prod_{d \geq 1} \sum_{k \geq 0} (-1)^k \binom{p_d}{k} t^{dk}.
\]
Isolating the coefficient of $t^{n}$, we obtain equation (ii).

Finally, we deduce (iii).  Beginning with Equation \eqref{eq:gf_relation} as above,
take the logarithm of each term and Taylor expand about $1$ to obtain
%
%giving
%\[
%-\log\left( 1 + \sum_{k \geq 1} h_k t^k \right)
%= \log\left( 1 - \sum_{m \geq 1} a_m t^m \right)
%= \sum_{d \geq 1} p_d \log( 1-t^d ).
%\]
%Replacing each logarithm with its Taylor expansion about $1$, this becomes
\[
 \sum_{r \geq 1} \frac{(-1)^{r-1}(\sum_{k \geq 1} h_k t^k )^r}{r}
= \sum_{r \geq 1} \frac{(\sum_{m \geq 1} a_m t^m )^r}{r}
= \sum_{d \geq 1} \sum_{j \geq 1} p_d \frac{t^{jd}}{j}.
\]
Now we isolate the coefficient of $t^{n}$ in each of the expressions: 
\[
\sum_{\beta \vDash n} \frac{(-1)^{\ell(\beta)-1} h_\beta}{\ell(\beta)}
= \sum_{\beta \vDash n} \frac{a_\beta}{\ell(\beta)}
= \frac{1}{n} \sum_{d | n} d p_{d} .
\]
Lastly, multiply the equation by $n$ and apply M\"{o}bius inversion to solve for $n p_{n}$.
\end{proof}

%\begin{remark}
Proposition \ref{prop:sequences} can be stated as biconditional theorem in the sense
that any one equality in the proposition implies, by reversing the calculations,
that the corresponding generating function relations in Equation~\eqref{eq:gf_relation}
will hold.
%\end{remark}
%In light of 

\begin{definition}
\label{defn:sequencetransfer}
The \emph{sequence transfer maps} relate sequences $(\vec{a}, \vec{h}, \vec{p})$ satisfying Equation~\eqref{eq:gf_relation}:
\begin{align*}
\begin{array}{rcl}
\phi_{p, a} : \CC^{\NN} & \to & \CC^{\NN} \\
\vec{p} & \mapsto & \vec{a}
\end{array},
\qquad
\begin{array}{rcl}
\phi_{h, a} : \CC^{\NN} & \to & \CC^{\NN} \\
\vec{h} & \mapsto & \vec{a}
\end{array},
\qquad\text{and}\qquad 
\begin{array}{rcl}
\phi_{p, h} : \CC^{\NN} & \to & \CC^{\NN} \\
\vec{p} & \mapsto & \vec{h}
\end{array}
\end{align*}
as well as their inverses
\[
\phi_{a, p} = \phi_{p, a}^{-1},
\qquad\qquad
\phi_{a, h} = \phi_{h, a}^{-1},
\qquad\qquad\text{and}\qquad\qquad
\phi_{h, p} = \phi_{p, h}^{-1}.
\]
\end{definition}

Following the conventions of \cite{OEIS}, the map $\phi_{p,h}$ is known
as the Euler transform of the sequence
% check A001970 = Euler transform applied twice to all 1's sequence = phi_ph(phi_ph([1,1,1,1,1,1,1,1,1,1]))
(and consequently $\phi_{h,p}$ is the inverse Euler transform).
Also following the conventions
of \cite{OEIS},
the map $\phi_{a,h}$ is known as the INVERT transform
% check A033453 = "INVERT" transform of squares A000290 = phi_ah([1,4,9,16,25,36,...])
and $\phi_{h,a}$ is the INVERTi transform.

\begin{example}
\label{ex:catalan}
Take $f(t) = 1 + t + 2 t^{2} + 5t^3 + 14t^4 + \cdots \in \CC[[t]]$,
so that $h_n$ is equal to the $n^{th}$ Catalan number
$\frac{1}{n+1}\binom{2n}{n}$.  It is well known that $f(t)$ satisfies the functional equation
\[
f(t) = \frac{1}{1 - t - t^2 - 2t^3 - 5t^4 - 14t^5 - \cdots} = \frac{1}{1-t f(t)}
\]
so we have that $\phi_{h,a}(1,2,5,14,42,132,\ldots) = \vec{a} = (1,1,2,5,14,42, \ldots)$
(that is, $a_n$ is the $n-1^{st}$ Catalan number).
We can use Proposition~\ref{prop:sequences} to calculate the first few values of $\vec{p} = (1,1,3,8,25, \ldots)$.
It follows from \cite[Section 5.1.3]{NT05} that $\vec{p}$ is equal to sequence \cite[A022553]{OEIS}.
%and $p_n$ is the number of binary Lyndon words of length $2n$
%with the same number of $0$s as $1$s (the definition of Lyndon word appears before
%Proposition \ref{prop:combinatorialinterpretation} below).
\end{example}

If the sequence $\vec{a}$ consists of entirely non-negative integers,
then $\vec{h} = \phi_{a, h}(\vec{a})$ and $\vec{p} = \phi_{a, p}(\vec{a})$ will also have combinatorial interpretation in terms of words.
Recall that a word in a set $X$ is a finite sequence $w = w_{1} w_{2} \ldots w_{\ell}$ of ``letters'' $w_{i} \in X$.  
If $X = \bigcup_{n \geq 1} X^{(n)}$ is a graded set, then define the \emph{degree} of a word to be
%$x \in X^{(n)}$ to be $n$, and let
\[
\mathsf{deg}(w_1 w_2 \ldots w_{\ell}) = \sum_{i = 1}^{\ell} \mathsf{deg}(w_{i})
\qquad\text{where $\mathsf{deg}(x) = n$ for all $x \in X^{(n)}$}.
%\qquad\text{for $d_{i}$ such that $w_{i} \in X^{(d_{i})}$}.
\]
For a fixed order on $X$, we order the words on $X$ lexicographically.  
The \emph{rotation} of a word $w = w_{1} w_{2} \ldots w_{\ell}$ is the word
\[
\mathsf{cyc}(w) =  w_{2} \ldots w_{\ell} w_{1}.
\]
This defines an operation of order $\ell$ on words of with $\ell$ letters.  
A word is \emph{Lyndon} if it is strictly smaller than each of $\mathsf{cyc}(w), \mathsf{cyc}^{2}(w), \ldots, \mathsf{cyc}^{\ell-1}(w)$.  
For instance, if $X = \{x < y\}$, then $xyxyy$ is a Lyndon word, but neither $xyxy$ nor $xyx$ are: $xyxy = \mathsf{cyc}^{2}(xyxy)$, while $xyx > xxy = \mathsf{cyc}^{2}(xyx)$.

If the sequence $\vec{a}$ is of non-negative integers rather than
any values from a field, there is a combinatorial interpretation for
$\phi_{a, h}(\vec{a})$ and $\phi_{a, p}(\vec{a})$.
The proposition below is well known, but the presentation of the
formulae in Proposition \ref{prop:sequences} does not match a reference that
we could find in the literature.  Hence we include the details of the
following result for completeness.

\begin{proposition}
\label{prop:combinatorialinterpretation}
Let $\vec{a} \in \NN^{\NN}$ so that there exists a graded set $X = \bigcup_{n \geq 1} X^{(n)}$ with $\veccard(X) = \vec{a}$.  
If $\vec{h} = \phi_{a, h}(\vec{a})$ and $\vec{p} = \phi_{a, p}(\vec{a})$, then for each $n \geq 1$, we have:
%Old/conflict version:
%Let $(\vec{h}, \vec{a}, \vec{p})$ be a triple of sequences in $\CC^{\NN}$ satisfying Equation \eqref{eq:gf_relation}.
%If $\vec{a}$ consists of only nonnegative integers, so that there exists a graded set $X = \bigcup_{n \geq 1} X^{(n)}$ with $|X^{(n)}| = a_{n}$ for each $n \geq 1$, we have:
\begin{enumerate}[itemsep = 0.5em]
\item $h_n$ is equal to the number of words of degree $n$ in the alphabet $X$ for all $n \ge 1$, and 

\item $p_n$ is equal to the number of Lyndon words of degree $n$ in the alphabet $X$ for all $n \ge 1$.
\end{enumerate}
\end{proposition}

\begin{proof}
To see (1), note that any degree $n$ word $w = w_1 w_2 \cdots w_{\ell}$ in $X$
defines a unique composition $(\mathsf{deg}(w_{1}), \mathsf{deg}(w_{2}), \ldots, \mathsf{deg}(w_{\ell}))$,
and for a particular composition $\beta \vDash n$,
every $w \in X^{(\beta_{1})} \times X^{(\beta_{2})} \times \cdots \times X^{(\beta_{\ell})}$ has this property.
Thus, the number of degree $n$ words in $X$ agrees with the formula for $h_{n}$ given in Proposition~\ref{prop:sequences} (i).  

For the second point the reader may refer to \cite[Theorem 4.9, Theorem 5.1]{Reutenauer-FreeLieAlgebras}
for a more complete exposition.

The proof will make use of the set
\[
I_{n} = \{ (i, w) \;|\; \text{degree $n$ words $w$ in $X$ and $1 \le i \le \mathsf{deg}(w_{1})$}\},
\]
which has size
$
|I_{n}| = \sum_{\beta \vDash n} \beta_{1} a_{\beta} .
$
Note that since $a_\beta = a_{\mathsf{cyc}^k(\beta)}$, then we have
\[
|I_{n}|
= \sum_{\beta \vDash n} \frac{1}{\ell(\beta)}
\sum_{k=1}^{\ell(\beta)} \beta_{1} a_{\mathsf{cyc}^{k}(\beta)}
= \sum_{\beta \vDash n} \frac{1}{\ell(\beta)}
\sum_{k=1}^{\ell(\beta)} \mathsf{cyc}^{-k}(\beta)_{1} a_{\beta}
= \sum_{\beta \vDash n} \frac{n a_{\beta}}{\ell(\beta)}~.
\]
We will show that 
\begin{equation}
\label{eq:LyndonMobius}
\sum_{\beta \vDash n} \frac{n a_{\beta}}{\ell(\beta)} = |I_{n}| = \sum_{d | n} d \; |\{\text{degree $d$ Lyndon words in $X$}\}|,
\end{equation}
from which M\"{o}bius inversion shows that the number of degree $n$ Lyndon words is equal to the formula for $p_{n}$ in terms of $\vec{a}$ given in Proposition~\ref{prop:sequences} (iii).

To start, we define a ``faux-cycling'' operation on $I_{n}$ by
\[
\mathsf{fcyc}(i, w) = \begin{cases} 
\big(i+1, w\big) & \text{if $i < \mathsf{deg}(w_{1})$} \\ 
\big(1, \mathsf{cyc}(w)\big) & \text{otherwise}.
\end{cases}
\]
For any degree $n$ word $w = w_{1} w_{2} \cdots w_{\ell}$ and  $1 \le a \le \ell$, 
\[
(0, \mathsf{cyc}^{a}(w)) = \mathsf{fcyc}^{\mathsf{deg}(w_{1}) + \cdots + \mathsf{deg}(w_{a-1})} (0, w).
\]
Thus, $\mathsf{fcyc}$ is periodic of order $n$.  Moreover, if $k$ is the minimal positive integer for which $\mathsf{cyc}^{k}(w) = w$, then the $\mathsf{fcyc}$ orbit of $(0, w)$---or $(i, w)$ for any $1 \le i \le \mathsf{deg}(w_{1})$---has size
\[
d = \mathsf{deg}(w_{1}w_{2}\cdots w_{k}).
\]
Finally, the cyclic shifts of $w' = w_{1}w_{2}\cdots w_{k}$ are distinct, so taking the unique minimal one, we obtain a Lyndon word of degree $d$.  Since $w$ is the $(\ell/k$)-fold concatenation of $w'$ with itself, the $d$-elements of the $\mathsf{fcyc}$-orbit of $(0, w)$ are uniquely determined by $w'$.
\end{proof}

\begin{example}\label{ex:balancedLyndon}
Continuing with Example \ref{ex:catalan}, we have that if
$h_n = a_{n+1} = \frac{1}{n+1}\binom{2n}{n}$, Proposition
\ref{prop:combinatorialinterpretation} says that $p_n$ is also equal to the number
of Lyndon words in an alphabet $X = \bigcup_{n\geq1} X^{(n)}$ with $a_n = |X^{(n)}|$.

However,~\cite[A022553]{OEIS} states that $p_n$ is is also equal to the number of balanced binary Lyndon words of size $n$.  A balanced binary word is a sequence $D = d_1d_2\ldots d_{2n} \in \{0, 1\}^{n}$ with an equal number of $1$s and $0$s, and we say that $D$ has size $n$.  We take the order $0 < 1$ when considering the Lyndon condition.  
We say that $D$ is a Dyck word of size $n$ if $\sum_{i=1}^k d_i \geq k/2$ for $1 \leq k \leq 2n$
and it is well known that the number of Dyck words of size $n$ is the Catalan number $\frac{1}{n+1}\binom{2n}{n}$.  

We have that $p_n$ is also the number of Lyndon words of the form
$(D_1, \ldots, D_r)$ where for $1 \leq i \leq r$,
$D_i$ is a Dyck word, including the empty word $\bullet$, and $r+|D_1|+\cdots+|D_r| = n$.  
If we order the set of Dyck words by size and then lexicographically within each size, then for $n = 8$ these Lyndon words are
$(111000)$, $(110100)$, $(101100)$, $(110010)$, $(101010)$,
$(\bullet, 1100)$, $(\bullet, 1010)$, $(\bullet, \bullet, 10)$.
From Example \ref{ex:catalan}, we confirm that $p_4 = 8$.
\end{example}

\begin{corollary}
If $\vec{a} \in \NN^{\NN}$, then $\vec{h} \geq \vec{p} \geq \vec{a} \geq \vec{0}$.
\end{corollary}
\begin{proof}
Let $X = \biguplus_{n \ge 1} X^{(n)}$ be a graded set with $|X^{(n)}| = a_{n}$.  
Considered as a word with one letter, each element $x \in X$ is a Lyndon word, so by Proposition~\ref{prop:combinatorialinterpretation} the inequalities above correspond to the set inclusions
\[
\{\text{degree $n$ words in $X$}\} 
\supseteq \{\text{degree $n$ Lyndon words in $X$}\} 
\supseteq X^{(n)}. \qedhere
\]
%We have that for each $i \geq 1$ and each $x \in X^{(i)}$ is a Lyndon word, hence $a_i \leq p_i$.
%Moreover each Lyndon word of degree $i$ is itself a word of degree $i$, so $h_i \geq p_i$.
\end{proof}

\begin{comment}
A \defn{set partition} $A$ of a finite set $X$ is a set of sets $A = \{ A_1, A_2, \ldots, A_k \}$ where $A_i \subseteq X$,
$A_i \cap A_j = \emptyset$ if $i \neq j$ and $\cup_{i=1}^k A_i = X$.  The sets $A_i$ are known as the \defn{parts}
of the set partition.  We indicate that $A$ is a set partition of
$[n] := \{1,2, \ldots, n\}$ by the notation $A \vdash [n]$ and the \defn{length} of the
set partition $A$ is $k$ and is denoted $\ell(A)$.

A set partition $A$ is called \defn{crossing} if there exists $A_i \neq A_j$ which
are parts of $A$ with $a < b < c < d$ and $a,c \in A_i$ and $b,d \in A_j$.  A set partition is said to be \defn{noncrossing} otherwise.
A set partition $A$ is said to be \defn{nesting} if there exists $A_i \neq A_j$ which
are parts of $A$ with $a < b < c < d$ and $a,d \in A_i$ and $b,c \in A_j$ and is \defn{nonnesting} otherwise.
\mike{The definition of non-nesting is probably not correct}

We will define a few combinatorial operations on set partitions that we will need to describe the algebra structure.
For $A \vdash [n]$ and $k \in \mathbb{Z}_{\geq 0}$, let $A_{\uparrow k}$ be the set partition
$\{ \{ a + k : a \in A_i \} : 1 \leq i \leq \ell(A) \}$.  If $A \vdash [n]$ is not equal to $B \cup C_{\uparrow k}$
for some set partitions $B \vdash [k]$ and $C \vdash [n-k]$ with $0 < k < n$, then we say that
$A$ is \defn{atomic}.

The algebra of symmetric functions in noncommuting variables was first considered by Wolf \cite{W36}
and much later in a more combinatorial setting by Rosas and Sagan \cite{RS06}.
We will define $\mathsf{NCSym}$ by its Hopf algebra structure as the linear span of set partitions.
We define the finite dimensional vector space
$\mathsf{NCSym}_n = \mathrm{span}_{\mathbb C}\{ \mathbf{h}_A : A \vdash [n] \}$
and set $\mathsf{NCSym} := \bigoplus_{n \geq 0} \mathsf{NCSym}_n$.

The product on $\mathsf{NCSym}$ is defined on the basis elements and extended linearly.
Let $A \vdash [k]$ and $B \vdash [n]$, then
$$\mathbf{h}_A \cdot \mathbf{h}_B = \mathbf{h}_{A \cup B_{\uparrow k}}~.$$

The coproduct on the Hopf algebra $\mathsf{NCSym}$ appears in \cite{B08}
and is equal to
$$\Delta(\mathbf{h}_A) = \sum_{A \subseteq [n]}
\mathbf{h}_{\mathrm{st}(A|_S)} \otimes \mathbf{h}_{\mathrm{st}(A|_{S^c})}~.$$
\mike{notation to be defined above: $A|_S$, $\mathrm{st}(A)$}
\end{comment}

\section{Sequences and universal enveloping algebras}

We now give the algebraic context for the results of Section~\ref{sec:SequencePrelims}.  
As in the introduction, we use the abbreviation \FGCCHA to mean \textsf{F}ree \textsf{G}raded \textsf{C}onnected \textsf{C}ocommutative \textsf{H}opf \textsf{A}lgebra, which we now formally define.

An \FGCCHA is a graded cocommutative Hopf algebra $H$ which is freely generated by some graded set $X = \biguplus_{n \ge 1} X^{(n)}$, meaning that as an algebra, $H$ is isomorphic to the tensor algebra
\[
\mathsf{T}(\CC X) = \bigoplus_{n \ge 0} \Big( \bigoplus_{k \ge 0} \CC X^{(k)}\Big)^{\otimes n}
\qquad\text{where}\qquad
\CC X^{(k)} = \CC\operatorname{-span}\{x \in X^{(k)}\},
\]
which is naturally graded so that all elements of $\CC X^{(k)}$ have degree $k$.  
Under this isomorphism, the coproduct of $H$ corresponds to a graded algebra homomorphism $\Delta: \mathsf{T}(\CC X) \to \mathsf{T}(\CC X) \otimes \mathsf{T}(\CC X)$, but it need not be the canonical coproduct of the tensor algebra.

Given any Hopf algebra $H$, the primitive elements of $H$ are elements of the graded subspace
\[
\mathcal{P}(H) = \{x \in H \;|\; \Delta(x) = x \otimes 1 + 1 \otimes x\}.
\]
The main result of this section relates the graded dimensions of an \FGCCHA to its generating set and primitives.  
This sort of result is well-known to experts, but we have not found a statement which incorporates grading in the literature.  

\begin{proposition}
\label{prop:SequencesAsHopfStructures}
Let $H$ be a \FGCCHA with generating set $X$.  The triple of sequences
\[
\left(\vecdim(H), \,
\veccard(X),\,
\vecdim\left(\mathcal{P}(H)\right) \right)
\]
satisfies Equation~\eqref{eq:gf_relation}.
\end{proposition}

The proof of Proposition~\ref{prop:SequencesAsHopfStructures} occupies the remainder of the section and makes use of the Cartier-Milnor-Moore and Poincare--Birkhoff--Witt theorems, which we state below.  First, however,  we give an example to illustrate the result.  

\begin{example}
The Hopf algebra of noncommutative symmetric functions $\mathsf{NSym}$ \cite{GKLLRT}
is the dual Hopf algebra to the quasisymmetric functions $\mathsf{QSym}$.
We take as a free generating set of this Hopf algebra
$X = \{ \mathbf{h}_1, \mathbf{h}_2, \mathbf{h}_3, \ldots \}$
where $\veccard(X) = (1,1,1,\ldots)$.
The basis of the degree $n$ homogeneous component are indexed by compositions
of $n$ and hence $\vecdim(\mathsf{NSym}) = (1,2,4,8,\ldots)$.
The dimensions of the primitives
of the the noncommutative symmetric functions
$\vecdim(\mathcal{P}(\mathsf{NSym})) = (1, 1, 2, 3, 6, 9, 18, 30, 56, \ldots)$
where the $n^{th}$ component of the sequence is equal to the number
of compositions of $n$ that are Lyndon \cite[Section 4]{H05}.
\end{example}

We now begin to assemble our proof of Proposition~\ref{prop:SequencesAsHopfStructures}.  
As stated above, this is a folklore result which we prove using well-known theorems for the sake of completeness.  

Given a graded Lie algebra $L = \bigoplus_{n \ge 1} L^{(n)}$, the universal enveloping algebra is the cocommutative Hopf algebra
\[
\mathcal{U}(L) = \mathsf{T}(L) \big/ \big\langle x\otimes y - y \otimes x - [x, y] \;|\; \text{$x, y \in L$} \big\rangle.
\]
There is a canonical cocommutative Hopf algebra structure on $\mathcal{U}(L)$ in which every element of $L$ is primitive, and by \cite[Theorem 1.4]{Reutenauer-FreeLieAlgebras} we have $\mathcal{P}(\mathcal{U}(L))=L$.  

The Poincar\'{e}--Brikhoff--Witt Theorem~\cite[??]{Reutenauer-FreeLieAlgebras} states for any given ordered homogeneous basis $\{z_i\}$ of $L$, the enveloping algebra $\mathcal{U}(L)$ has a homogeneous basis 
\[
\{ z_{i_1}z_{i_2}\cdots z_{i_k} \;|\; i_1 \leq i_2 \leq \cdots \leq i_k\}.
\]
This statement carries important information about the dimension of $\mathcal{U}(L)$, namely that 
\[
\vecdim(\mathcal{U}(L))_{n}= \#\{\text{multisubsets $S$ of $\{x_{i}\}$} \;|\; \sum_{z \in S} m_{z}(S) \mathsf{deg}(x) = n\},
\]
where $m_{z}(S)$ denotes the multiplicity of $z$ in $S$.

For any \FGCCHA $H$, the space $\mathcal{P}(H)$ is a Lie algebra under the bracket $[x, y] = xy - yx$.  
The Cartier-Milnor-Moore theorem~\cite[see Theorem 5.18]{MM65} states there is a Hopf algebra isomorphism 
\[
\begin{array}{rcl}
\mathcal{U}(\mathcal{P}(H)) &\to& H \\
\mathcal{P}(H) \ni x & \mapsto&  x
\end{array}
\]

\begin{proof}[Proof of Proposition~\ref{prop:SequencesAsHopfStructures}]
Let $\vec{h} = \vecdim(H)$, $\vec{a} = \veccard(X)$, and $\vec{p} = \vecdim(\mathcal{P}(H))$.  We will show that $\vec{h} = \phi_{a, h}(\vec{a})$ and $\vec{p} = \phi_{h, p}(\vec{p})$, from which the claim follows.  

For $n \ge 1$, $h_{n}$ is the number of degree $n$ monomials in $X$, which is also the number of degree $n$ words in $X$.  
Applying Proposition~\ref{prop:combinatorialinterpretation}, $\vec{h} = \phi_{a, h}(\vec{a})$.

On the other hand, the Milnor--Moore theorem states that $H$ is
isomorphic to the universal enveloping algebra of $\mathcal{P}(H)$.
By the Poincare--Birkhoff--Witt theorem, for any fixed homogeneous basis
$Y = \biguplus_{n \ge 0} Y^{(n)}$ of $\mathcal{P}(H)$, $h_{n}$ counts
the multisets of $Y$ whose elements have degree summing (with repetition)
to $n$.  Since $|Y^{(n)}| = p_{n}$ for all $n \ge 1$, the number of these
multisets is exactly
\[
\sum_{\lambda \vdash n} \prod_{d \geq 1} \binom{p_d + m_d(\lambda) -1}{m_d(\lambda)},
\]
so by Proposition~\ref{prop:sequences} we have $\vec{p} = \phi_{h, p}(\vec{h})$.
\end{proof}

%\subsection{Free algebras}
%\label{sec:freealgebras}
%
%In this section we define the free algebra on a graded set and state its universal property.  
%Let $X$ be a graded set, 
%\[
%X = \biguplus_{n \ge 1} X_{n}.
%\]
%The free graded algebra on the set $X$ is\mike{what are the $g_x$?}
%\[
%\mathcal{F}(X) = \bigoplus_{n \ge 0} \bigoplus_{\alpha \vDash n} 
%\CC\operatorname{-span}\{g_{x_{i_{1}} }g_{x_{i_{2}}} \cdots g_{x_{i_{\ell}}} \;|\; x_{i_{k}} \in X_{\alpha_{k}}\},
%\]
%where the degree zero component is $\CC 1_{\mathcal{F}(X)}$.  This algebra $\mathcal{F}(X)$
%has the following \emph{universal property}\mike{reference?}:
%
%\begin{property}
%\label{prop:universalfree}
%Let $A = \bigoplus_{n \ge 0} A_{i}$ be a graded algebra.  For every function $\phi: X \to A$ such that $\phi(X_{i}) \subseteq A_{i}$, there exists a unique graded algebra homomorphism $\tilde{\phi}: \mathcal{F}(X) \to A$ such that $\tilde{\phi}|_{X} = \phi$, i.e.
%\[
%    \begin{tikzpicture}
%        \node at (0, 0) (X) {$X$};
%        \node at (3, 0) (F) {$A$};
%        \node at (0, -1) (A) {$\mathcal{F}(X)$};
%        \draw[->] (X) -- node[above] {$\phi$} (F);
%        \draw[->] (X) -- node[left] {inclusion} (A);
%        \draw[->, dashed] (A) -- node[below right] {$\tilde{\phi}$} (F);
%        \end{tikzpicture}.
%\]
%\end{property}
%
%
%\begin{proposition}
%\label{prop: free algebra hilbert series}
%Given a sequence $\vec{a} = (a_{1}, a_{2}, \ldots)$ of nonnegative integers, let $X$ be a graded set with $|X_{i}| = a_{i}$, and consider the free algebra $\mathcal{F}(X)$.  Then the Hilbert series for $\mathcal{F}(X)$ is
%\[
%\mathrm{Hilb}(\mathcal{F}(X); t) =  \frac{1}{1 - a_{1}t - a_{2}t^{2} - a_{3}t^{3} - \cdots}.
%\]
%%
%%
%%
%%Given a sequence $\vec{a} = (a_{1}, a_{2}, \ldots)$ of nonnegative integers, let $\mathfrak{L}(\vec{a})$ denote the free, graded Lie algebra with $a_{i}$ generators in degree $i$, and zero generators in degree zero.  The universal enveloping algebra
%%\[
%%\mathcal{U}\big(\mathfrak{L}(\vec{a})\big)
%%\]
%%
%%\begin{enumerate}
%%
%%\item is a free, cocommutative, graded, connected Hopf algebra.
%%
%%\item has Hilbert series
%%
%%\[
%%f(t)=\frac{1}{1 - a_{1}t - a_{2}t^{2} - a_{3}t^{3} - \cdots}.
%%\]
%%
%%\end{enumerate}
%\end{proposition}
%
%\begin{proof}
%%
%%\begin{enumerate}
%%
%%    \item From Jacobson \cite[more specific reference]{jacobson2013lie}\lucas{This is now proved in section~\ref{sec:envelopingalgebras}, we should refer to this or remove this claim entirely.}
%%
%%    \item 
%    Expand the following generating function as a geometric series, we have 
%
%    $${\begin{split}
%        \frac{1}{1 - a_{1}t - a_{2}t^{2} - a_{3}t^{3} - \cdots} &= \sum_{n=0}^\infty (a_1t+a_2t^2+\cdots)^n~.
%    \end{split}}$$
%
%    Expanding this sum we have the following equality
%
%    $$\frac{1}{1 - a_{1}t - a_{2}t^{2} - a_{3}t^{3} - \cdots}
%    =\sum_{n=0}^\infty \left(\sum_ {\beta \vDash n} a_\beta \right) t^n.$$
%
%    Given $\Vec{a} = (a_{1}, a_{2}, \ldots)$, we can consider that $\mathfrak{L}(\Vec{a})$\mike{what is $\mathfrak{L}(\Vec{a})$?  Should this be
%    $\mathcal{F}(X)$?}
%is generated by the set
%    $$X=\biguplus_{i \geq 1}\{x_{(i,j)}: 1 \leq j \leq a_i\}$$
%     where for each $j$, $x_{(i,j)}$ is a generator of degree $i$.\mike{what do we mean ``is a generator''? of what?}
%
%From ~\cite[Theorem 7 (Witt)]{jacobson2013lie} we have that a basis for the graded components $\mathcal{U}\big(\mathfrak{L}(\vec{a})\big)_n$ is the set of monomials of the form $m=x_{(\beta_1,j_1)} \cdots x_{(\beta_\ell,j_\ell)}$ where
%     for each $d$, $1 \leq j_d \leq a_{\beta_{d}}$ with $\beta=(\beta_1,\dots,\beta_\ell)$ is a composition of $n$.
%By counting the number of monomials of degree $n$,
%we conclude that the dimension of each graded component of $\mathcal{F}(X)$
%is given by $\sum_ {\beta \models n} a_\beta$
%which is the coefficient of $t^n$ in the generating function.
%%\end{enumerate}
%
%\end{proof}
%
%\subsection{Universal Enveloping Algebras}
%\label{sec:envelopingalgebras}
%
%Let $L$ be a graded Lie algebra with bracket $[\cdot, \cdot]$.  The universal enveloping algebra is the graded associative algebra on the space
%\[
%\mathcal{U}(L) = \bigoplus_{n \ge 0} \bigoplus_{\alpha \vDash n} L_{\alpha_{1}} \otimes L_{\alpha_{2}} \otimes \cdots \otimes L_{\alpha_{\ell}},
%\]
%where the degree zero component is $\CC 1_{\mathcal{U}(L)}$.
%Multiplication in $\mathcal{U}(L)$ is concatenation of tensors,
%subject to the relations of the form $a \otimes b - b \otimes a = [a, b]$
%for all $a, b \in L$.  This algebra $\mathcal{U}(L)$ has the following \emph{universal property}:\mike{reference?}
%
%\begin{property}
%\label{prop:universalLie}
%Let $A = \bigoplus_{n \ge 0} A_{i}$ be a graded algebra.  For every linear map $\psi: L \to A$ such that $\psi(L_{i}) \subseteq A_{i}$ and
%\[
%\psi([a, b]) = \psi(a)\psi(b) - \psi(b)\psi(a)
%\qquad\text{for all $a, b \in L$},
%\]
%there exists a unique graded algebra homomorphism $\tilde{\psi}: \mathcal{U}(L) \to A$ such that $\tilde{\psi}|_{L} = \psi$, i.e.
%\[
%    \begin{tikzpicture}
%        \node at (0, 0) (L) {$L$};
%        \node at (3, 0) (U) {$A$};
%        \node at (0, -1) (A) {$\mathcal{U}(L)$};
%        \draw[->] (X) -- node[above] {$\psi$} (U);
%        \draw[->] (X) -- node[left] {inclusion} (A);
%        \draw[->, dashed] (A) -- node[below right] {$\tilde{\psi}$} (U);
%        \end{tikzpicture}.
%\]
%\end{property}
%
%Given a graded set $X = \biguplus_{i \ge 1} X_{i}$, we also define the free Lie algebra on $X$ to be the graded Lie algebra
%\mike{what are the $h_x$?}
%\[
%\mathfrak{L}_{X} = \bigoplus_{n \ge 1} \bigoplus_{\alpha \vDash n} \CC\operatorname{-spanning}\{\text{brackets of $h_{x_{i_{1}}}, \ldots, h_{x_{i_{\ell}}}$} \;|\; x_{i_{k}} \in X_{\alpha_{k}} \},
%\]
%subject only to anticommutativity and the Jacobi identities.\mike{what are these relations?}
%
%%The following is quoted word for word from \cite{jacobson2013lie}. Lie algebras arise from associative algebras in a very simple way. Let $A$ be an associative algebra. If $x,y,\in A$, then we define the \textit{Lie product} or (additive) \textit{commutator} of $x$ and $y$ as
%%
%%$$[x,y]=xy-yx.$$
%%
%%One checks immediately that
%%
%%$$\begin{matrix}
%%    [x_1+x_2,y]=[x_1,y]+[x_2,y], \\
%%    [x,y_1+y_2]=[x,y_1]+[x,y_2],\\
%%    \alpha[x,y]=[\alpha x,y]=[x,\alpha y].
%%\end{matrix}$$
%%
%%Thus the products $[x,y]$ satisfies all the conditions on the product in a Lie algebra. The Lie algebra obtained in this way is called \textit{the Lie algebra} of the associative algebra $\mathfrak{U}$. We shall denote this Lie algebra as $\mathfrak{U}_L$.
%
%\begin{thm}[Theorem $7$ (Witt),\cite{jacobson2013lie}]
%Let $X$ be an arbitrary graded set.  Then
%\[
%\mathcal{U}(\mathfrak{L}_{X}) \cong \mathcal{F}(X).
%\]
%\label{thm: free UEA}
%%
%%Let
%%
%%
%%Let $X$ be an aribtrary set and let $\mathcal{F}$ denote the free algebra (freely) generated by $X$. Let $\mathfrak{F}^{\mathfrak{L}}$ denote the subalgebra of $\mathfrak{F}_L$ generated by the elements of $X$. Then $\mathfrak{F}^{\mathfrak{L}}$ is a free Lie algebra generated by $X$ and $\mathfrak{F}$ is the universal enveloping algebra of $\mathfrak{F}^{\mathfrak{L}}$. 
%\end{thm}
%\begin{proof}
%    Let \(X = \biguplus_{n \ge 1} X_{n}\) be a graded set and let \(\mathcal{F}(X)\) be the free algebra generated by \(X\). Also, let \( \mathcal{U} (\mathfrak{L}_{X})\) be the Universal Enveloping algebra of the free Lie algebra generated by \(X\). 
%    Define the map 
%    \begin{align*}
%        \begin{array}{rcl}
%            \phi :  X \longrightarrow & \mathcal{U} (\mathfrak{L}_{X}) \\
%             x  \longmapsto  & h_x \\ 
%        \end{array}
%    \end{align*}
%    where \(h_x =  h_{x_{i_{1}} }h_{x_{i_{2}} }\cdots h_{x_{i_{k}} } \in \mathcal{U} (\mathfrak{L}_{X})\) is the product of formal generators in \(X\). 
%    By Property~\ref{prop:universalfree} we have the unique graded homomorphism 
%    \begin{align*}
%        \begin{array}{rcl}
%            \tilde{\phi}: \mathcal{F}(X) \longrightarrow & \mathcal{U} (\mathfrak{L}_{X}) \\
%            g_x \longmapsto & h_x
%        \end{array}
%    \end{align*}
%    where \(g_x = g_{x_{i_{1}} }g_{x_{i_{2}} }\cdots g_{x_{i_{\ell}} } \in \mathcal{F}(X)\) is also the product of formal generators in \(X\). Similarly, 
%    define 
%    \begin{align*}
%        \begin{array}{rcl}
%            \psi :  \mathfrak{L}_{X} \longrightarrow &\mathcal{F}(X) \\
%        x  \longmapsto  & g_x .\\
%        \end{array}
%    \end{align*}
%    From Property~\ref{prop:universalLie} we obtain another unique graded homomorphism 
%    \begin{align*}
%        \begin{array}{rcl}
%            \tilde{\psi}: \mathcal{U}(\mathfrak{L}_{X}) \longrightarrow & \mathcal{F} (X) \\
%            h_x \longmapsto & g_x.
%        \end{array}
%    \end{align*}
%    To see that \(\mathcal{U}(\mathfrak{L}_{X}) \cong \mathcal{F}(X)\), notice that
%    \begin{align*}
%        \tilde{\phi}(\tilde{\psi}(h_{x_{i_{1}} }h_{x_{i_{2}}} \cdots h_{x_{i_{k}}}))  
%        &= \tilde{\phi}(\tilde{\psi}(h_{x_{i_{1}} })\tilde{\phi}(\tilde{\psi}(h_{x_{i_{2}} } )\cdots \tilde{\phi}(\tilde{\psi}(h_{x_{i_{k}} })) \\
%        &=  \tilde{\phi}(g_{x_{i_{1}} })\tilde{\phi}(g_{x_{i_{2}} })\cdots \tilde{\phi}(g_{x_{i_{\ell}} })\\
%        &= h_{x_{i_{1}} }h_{x_{i_{2}} } \cdots h_{x_{i_{k}} }
%    \end{align*} 
%    Similarly, it can also be shown that \(\tilde{\psi} \circ \tilde{\phi} (g_x) = g_x\) and as such \(\tilde{\phi} \circ \tilde{\psi} = \tilde{\psi} \circ \tilde{\phi} = id\).
%    Therefore, since \(\tilde{\phi}\) and \(\tilde{\psi}\) are inverses of each other,  \(\mathcal{U}(\mathfrak{L}_{X}) \cong \mathcal{F}(X)\).
%\end{proof}

%Let $\Phi$ be a field. For the sake of simplcity we shall now restrict our attention to the case of a finite set set $X=\{x_1,x_2,\cdots,x_r\}$. Then
%define
%$$\mathfrak{M}:=\Phi x_1 \oplus \Phi x_2 \oplus \cdots \oplus \Phi x_r$$
%
%and $\mathfrak{F}= \Phi 1 \oplus \mathfrak{M} \oplus (\mathfrak{M} \otimes \mathfrak{M}) \oplus \cdots$, and we write $\mathfrak{F}=\Phi\{x_1,\ldots,x_r\}$. The algebra $\mathfrak{F}$ is graded with $\mathfrak{M}_m=\mathfrak{M}\otimes\mathfrak{M}\otimes\cdots \otimes \mathfrak{M}$ ($m$-times) as the space of the homogeneous elements of degree $m$. A basis for this space is the set of monomials of the form $x_{i_1}x_{i_2}\cdots x_{i_m}$, $i_j=1,2,\cdots,r$; hence $\dim \mathfrak{M}_m=r^m$.

\section{Characterization of \FGCCHAs by graded dimension}

This section gives a classification of isomorphism types of \FGCCHAs via the associated integer sequences.  
This begins with Aliniaeifard and Thiem's result.  

\begin{thm}[{\cite[Theorem 4.2]{AT22}}]
\label{thm:AT}
Let $G$ and $H$ be \FGCCHAs.  Then $H \cong G$ if and only if $\vecdim(H) = \vecdim(G)$.
\end{thm}

What remains is to determine which sequences $\vec{h} \in \CC^{\NN}$ appear as the graded dimension of an \FGCCHA, a novel result.  In fact, we go further by identifying an explicit representative of each isomorphism class of \FGCCHA.  

\begin{definition}[{see~\cite[Section 0.2]{Reutenauer-FreeLieAlgebras}}]
For any sequence $\vec{a} = (a_{1}, a_{2}, \ldots)$ of nonnegative integers,
let $\mathfrak{L}(\vec{a})$ be the free graded Lie algebra with
$a_{n}$ generators in degree $n \ge 1$.
\end{definition}
\lucas{Should just define as tensor algebra?  Lucas will TODO}

We now state our classification; recall the map $\phi_{h, a}$ from Definition~\ref{defn:sequencetransfer} and the universal enveloping algebra map $\mathcal{U}(-)$ defined in Section~\ref{}.

\begin{thm}
\mike{TODO: I've changed the statement of the Theorem so that it is characterized by $\vec{a}$ sequences, but not the proof of Theorem 2.}
\label{thm:a sequence}
There is a bijection
\[
\begin{array}{rcl}
\left\{ \text{sequences }\vec{a} \in \NN^{\NN} \right\} & \to & \left\{ \begin{array}{c} \text{Isomorphism classes} \\ \text{of \FGCCHA} \end{array} \right\} \\[2ex]
\vec{a} & \mapsto & \mathcal{U}(\mathfrak{L}(\vec{a}))) \\
\phi_{h,a}(\vecdim(H)) & \mapsfrom & H
\end{array}.
\]
\end{thm}

This result is proved later in the section; we continue our discussion after the following example.

\begin{example}
Many combinatorially interesting sequences correspond to \FGCCHAs, but it is not obvious when this is the case; we illustrate this by considering two closely-related sequences.
\begin{enumerate}
\item $(1,1,2,3,5,8,13,\ldots)$, the Fibonacci sequence

\item $(2,1,3,4,7,11,18,\ldots)$, the Lucas sequence
\end{enumerate}
Let us consider the Fibonacci sequence $\vec{f} = (1,1,2,3,5,8,13,\ldots)$
as if it could be the graded dimensions of some Hopf algebra.
We compute first that $\phi_{h,a}(\vec{f}) = (1,0,1,0,1,0,\ldots)$
and by Theorem \ref{thm:a sequence} conclude
that there is a Hopf algebra $H$ with a basis indexed by
compositions with only odd parts and with one generator at each odd degree.
This Hopf algebra $H$ has graded dimension equal to $\vecdim(H) = \vec{f}$.

Next consider the Lucas sequence $\vec{\ell} = (2,1,3,4,7,11,18,\ldots)$
as if it could also be the graded dimensions of a Hopf algebra.  We compute that $\phi_{h,a}(\vec{\ell})$ and determine
that the sequence is alternating and begins $(2, -3, 7, -13,\ldots)$ and so we must conclude
that there does not exist a Hopf algebra with graded dimensions equal to $\vec{\ell}$.
\end{example}

The remainder of the section concerns the technical question of explicitly realizing the Aliniaeifard--Thiem isomorphism in every possible way; this has further applications in later sections.
Given an \FGCCHA $H$, and \emph{ordered primitive generating set} with be a sequence of tuples 
\[
\vec{X} = \Big(\vec{X}^{(n)} = (x_{1}^{(n)}, x_{2}^{(n)}, \ldots, x_{a_{n}}^{(n)})\Big)_{n = 1}^{\infty}
\qquad\text{with}\qquad
\begin{array}{l}
%\text{(1) $\vec{X}^{(n)} = (x_{1}^{(n)}, x_{2}^{(n)}, \ldots, x_{1}^{(n)})$} \\
\text{(1) each $x_{i}^{(n)} \in \mathcal{P}(H)_{n}$ , and} \\
\text{(2) $H$ is freely generated by $\vec{X}$.}
\end{array}
\]
Let
\[
\OPG(H) = \left\{ \text{ordered primitive generating sets of $H$} \right\}.
\]
The second main result of the section is the the following.

\begin{thm}
\label{thm:OPGmaps}
Let $H$ be an \FGCCHA with $\vec{h} = \vecdim(H)$.  Then writing $\vec{a} = \phi_{h, a}(\vec{h})$ and $\vec{p} = \phi_{h, p}(\vec{h})$, we have bijections
\[
\begin{array}{rcl}
\{\text{isomorphisms $\mathcal{U}(\mathfrak{L}(\vec{a})) \to H$}\} &\leftrightarrow& \OPG(H) \\
\phi & \mapsto & \phi(X) \\
\end{array}
\]
and
\[
\begin{array}{rcl}
\OPG(H) &\leftrightarrow& \left\{ \begin{array}{c}
\text{Sequences $\big(A^{(n)} \in \mathrm{Mat}_{a_{n} \times p_{n}}(\CC)\big)_{n=1}^{\infty}$ } \\
\text{with $\det\big( (A^{(n)}_{i, j})_{1 \le i, j \le a_{n}} \big) \neq 0$ for $n \ge 1$}
\end{array}\right\}
\end{array}.
\]
\end{thm}

\begin{example}
\label{ex:OPGexample}
\lucas{TODO: Clean this example up}
We do an example for the Catalan \FGCCHA.  We take $\mathcal{U}(\mathfrak{L}(\vec{a}))$ to be generated by Dyck words (of one smaller degree) and $H$ to be the noncrossing subalgebra (in some basis) of $\mathsf{NCSym}$.
By (Lauve--Mastanek \cite{LM11} or Bergeron et.~al.), we have a set or primitive
generators for $\mathsf{NCSym}$, which gives an isomorphism by mapping Dyck words to primitives
\end{example}

\subsection{Proof of Theorem~\ref{thm:a sequence}}

We will make use the following established results
about free Lie algebras from~\cite{Reutenauer-FreeLieAlgebras}.
\mike{TODO: I changed the statement of the theorem, but not the text
of the proof, make sure that these align}

\begin{proposition}[{\cite[??]{Reutenauer-FreeLieAlgebras}}]
\label{prop:LieToFree}
\lucas{TODO: find exact citation for this}
For an sequence $\vec{a}$ of a nonnegative integers, the enveloping algebra $\mathcal{U}(\mathfrak{L}(\vec{a}))$ is an \FGCCHA which is generated in degree $n$ by the $a_{n}$-many primitive generators of $\mathfrak{L}(\vec{a})$.
\end{proposition}

We now prove Theorem~\ref{thm:a sequence}; recall Proposition~\ref{prop:SequencesAsHopfStructures}.

\begin{proof}[Proof of Theorem~\ref{thm:a sequence}]
Given a sequence $\vec{h}$ for which $\phi_{h, a}(\vec{h})$ consists of nonnegative integers, Proposition~\ref{prop:LieToFree} states that $\mathcal{U}(\mathfrak{L}(\phi_{h, a}(\vec{h})))$ is an \FGCCHA freely generated by a set with graded cardinality $\phi_{h, a}(\vec{h})$, so the given map is well-defined.  

Now consider injectivity.  By Proposition~\ref{prop:SequencesAsHopfStructures},
\[
\vecdim\big(\mathcal{U}(\mathfrak{L}(\phi_{h, a}(\vec{h})))\big) = \phi_{a, h}\big(\phi_{h, a}(\vec{h})) = \vec{h},
\]
which by Aliniaeifard and Thiem's result (Theorem~\ref{thm:AT}) is an invariant of the isomorphism type.  Thus, for any two sequences $\vec{h}, \vec{g}$ in the domain, 
\[
\mathcal{U}(\mathfrak{L}(\phi_{h, a}(\vec{h}))) \cong \mathcal{U}(\mathfrak{L}(\phi_{h, a}(\vec{g})))
\qquad\text{if and only if}\qquad
\vec{h} = \vec{g}.
\]

Finally for surjectivity, fix an isomophism class of \FGCCHA and let $H$ be a representative of that class.  Then the sequence $\vec{h} = \vecdim(H)$ belongs to the domain, since by  Proposition~\ref{prop:SequencesAsHopfStructures} $\phi_{h, a}(\vec{h})$ is the graded cardinality of its free generating set.  
Moreover, we claim that our isomorphism class is the image of $\vec{h}$.  
As we deduced in the previous paragraph, the dimension sequence of $\mathcal{U}(\mathfrak{L}(\phi_{h, a}(\vec{h})))$ is $\vec{h}$, which by definition is the dimension sequence of $H$.  
Thus, by Aliniaeifard and Theim (Theorem~\ref{thm:AT}), $H \cong \mathcal{U}(\mathfrak{L}(\phi_{h, a}(\vec{h})))$
\end{proof}

We conclude by giving an alternative sequence-based characterization of \FGCCHAs which follows as a Corollary of Theorem~\ref{thm:a sequence}.\lucas{We will fold this into a remark after theorem 2.}

\begin{corollary}
\label{prop:UniqueBySequence}
There is a bijection
\[
\begin{array}{rcl}
\left\{ \begin{array}{c} \text{sequences $\vec{p} \in \CC^{\NN}$ for which $\phi_{p,a}(\vec{p})$} \\ \text{contains only nonnegative integers} \end{array} \right\} & \to & \left\{ \begin{array}{c} \text{Isomorphism classes} \\ \text{of \FGCCHA} \end{array} \right\} \\[2ex]
\vec{h} & \mapsto & \mathcal{U}(\mathfrak{L}(\phi_{p,a}(\vec{p}))) \\
\vecdim\big(\mathcal{P}(H)\big) & \mapsfrom & H
\end{array}.
\]
%	Given a sequence of nonnegative integers $\vec{p} = (p_0, p_1, p_2, \ldots)$, there exists a \FGCCHA $H$ with $\vecdim(\mathcal{P}(H)) = \vec{p}$ if and only if $\phi_{p,a} \geq 0$.
\end{corollary}

\begin{proof}	
    Let \(\vec{p} = (p_0, p_1, p_2, \ldots)\) be a nonnegative integer sequence and let \(H\) be a 
    free graded connected cocommutative Hopf algebra such that \(\dim \mathcal{P}(H)_n = p_n\).
    By Proposition~\ref{prop:sequences} we have that
    \[
        h_n = \sum_{\lambda \vdash n} \prod_{d \geq 1} \binom{p_d + m_d(\lambda) -1}{m_d(\lambda)}
    \]
    so that \(\vec{p}\)  corresponds to a sequence \(\vec{h} = (h_0, h_1, h_2, \ldots)\) which is clearly nonnegative. 
    Hence \(a_n \geq 0\) 
    for all \(n \geq 0\) by Theorem~\ref{thm:a sequence}.

    Supposing \(a_n \geq 0\) for all \(n \geq 0\), then by Theorem~\ref{thm:a sequence} we have that \(\vec{a}\)
    generates a free graded connected
    cocommutative Hopf algebra \(H\), and we let \(p_n\) count the number of primitive elements of \(H\) at degree \(n\).
    Hence \(\vecdim(\mathcal{P}(H)_n) = \vec{p}\geq 0\).

    %Suppose that $p_n$ correspond to the dimension of a free cocommutative graded connected Hopf algebra $H$ at degree $n$. 
    %By ~\cite[Theorem 2.2]{KK95} and after manipulation we get 
    %$a_n = \sum_{\lambda \vdash n} (-1)^{\ell(\lambda)-1} \prod_{d \geq 1} \binom{p_d}{m_d(\lambda)}$ \felix{Is this enough?}.
    % By freeness, there exist graded sets X generating $H$. Hence, by Lemma $3.4$, we have that  there exists a set of primitives 
    %$X'=\biguplus_{n \geq 1}X_n'$ that freely generates $H$. We have that the $a_n$ are nonnegative integers since they correspond 
    %to the number of free generators at degree $n$ of $H\cong \mathcal{U}(\mathfrak{L}_{X'})$. Suppose that 
    %$a_n = \sum_{\lambda \vdash n} (-1)^{\ell(\lambda)-1} \prod_{d \geq 1} \binom{p_d}{m_d(\lambda)}$ such that $a_n\geq 0$, 
    %by Theorem $3$, we have that $\mathfrak{L}(\mathbf{a})\cong \mathcal{P}(H)$. Hence, we have that $\dim(\mathfrak{L}(\mathbf{a})=\mathbf{p}$. 
    %We have that the $a_n$ are nonnegative integers since they correspond to the dimension of the free Lie aglebra $\mathfrak{L}(\mathbf{a})$.
    %\felix{Really not sure about this part.} 
    %The relationship between $\mathbf{p}$ and $\mathbf{a}$ followed by the construction of $\mathcal{F}(X')\cong\mathcal{U}(\mathfrak{L}_{X'})\cong H$ 
    %given by Theorem $1$ and Lemma $3.4$  gives us a way to construct the Hopf algebra associated to $\mathbf{p}$.
\end{proof} 


\subsection{Ordered primitive generators and Theorem~\ref{thm:OPGmaps}}

We now consider the problem of explicitly constructing isomorphisms between \FGCCHAs via the set $\OPG(H)$ defined at the beginning of the section.  We will make use of the \emph{Eulerian Idempotent}\lucas{TODO: Add original citation or reference (text) for eulerian idempotent.  First place to look: Aguiar--Lauve's ``Adams operator'' paper.}
\[
\mathbf{e}(x) = x - \frac{1}{2}\mu \circ \Delta_{+}(x) + \frac{1}{3} \mu^{(2)} \circ \Delta_{+}^{(2)}(x) - \cdots 
\]
where $\Delta_{+}$ denotes the positive coproduct, $\Delta_{+}(x) = \Delta(x) - x \otimes 1 - 1 \otimes x$, and we write $\mu^{(k)}$ and $\Delta^{(k)}$ for the $k$th composition of the multiplication and comuliplication maps, respectively.  (By associativity and coassociativity, these are well defined).

\lucas{TODO: Explain how the next result fits into the literature.  I seem to remember it being a novel statement, and that e.g. Aguiar Mastanek actually don't seem to know that it holds in general?}

\begin{lemma}
\label{lemma:primitive generators}
Let $H$ be a graded connected cocommutative Hopf algebra which is freely generated by a graded set $X = \biguplus_{n \geq 1} X_{n}$.  
Then the graded set
\[
\mathbf{e}(X) = \biguplus_{n \ge 1} \{\mathbf{e}(x) \;|\; x \in X_{n}\}
\]
comprises a complete set of primitive generators of $H$.
\end{lemma}

\begin{proof}
\lucas{TODO:Distinguish between statements that are (1) obvious upon reflection, (2) require justification/proof, and (3) require a reference.  Anything that isn't (1) should have a written explanation/citation.}
        We will first show $\mathbf{e}(X)$ is algebraically independent in $H$. Fix a total order in each $X_n$, and extend to a total order on $X$, such that if $x\in X_i, y\in X_j$ and $i>j$, then $x<y$. For any $x\in X$, we have
        \[
        \mathbf{e}(x) = x + f((x_i)),
        \]
        where $f((x_i))$ is some polynomial in generators $x_i$ of lesser degree than $x$.
        Then the minimal\lucas{``Minimal term'' needs a definition/explanation.  Can we just prove this modulo $H^{2}$ or something similar?} term in $\mathbf{e}(x)$ is $x$.
        Moreover, for any polynomial $g$, $g(\mathbf{e}(x_1),...,\mathbf{e}(x_n))$ has the same minimal term as $g(x_1,...,x_n)$, and therefore $g(\mathbf{e}(x_1), \ldots, \mathbf{e}(x_n)) = 0$ if and only if $g(x_1, \ldots, x_n) = 0$.
        Therefore $\mathbf{e}(X)$ is an algebraically independent set.
    
        To show $\mathbf{e}(X)$ generates $H$, observe that in each degree $n$, the monomials in $X$ of degree $n$ form a basis for $H_n$. The degree $n$ monomials in $\mathbf{e}(X)$ are linearly independent, because the $\mathbf{e}(X)$ are algebraically independent. Because every monomial $\mathbf{e}(x_1)\mathbf{e}(x_2)\cdots \mathbf{e}(x_k)$ corresponds uniquely to the monomial $x_1x_2 \cdots x_k$, the degree $n$ monomials in $\mathbf{e}(X)$ form a basis of $H_n$, hence $\mathbf{e}(X)$ freely generates $H$.\lucas{I'm not sure this is quite correct, at least as written.}
\end{proof}

\begin{proof}[Proof of Theorem \ref{thm:OPGmaps}]
    Let us fix $X = \biguplus_{n \geq 1} \vec{X}^{(n)} \in \OPG(\mathcal{U}(\mathfrak{L}(\vec{a}))$. For a graded Hopf algebra isomorphism $\phi:\mathcal{U}(\mathfrak{L}(\vec{a})) \to H$, it is clear that $\phi(X)$ is a set of free homogeneous generators for $H$. We may induce an ordering on $\phi(X)$ from the ordering on $X$, so that 
    \[
    \vv{\phi(X)}^{(n)} = \left(\phi\left(x_1^{(n)}\right), \phi\left(x_2^{(n)}\right), \ldots, \phi\left(x_{a_n}^{(n)}\right)\right),
    \]
    hence $\phi(X)$ is a uniquely determined member of $\OPG(H)$.

    The assignment $\phi \mapsto \phi(X)$ is injective, because any two \FGCCHA isomorphisms equal on a set of generators extend to the same isomorphism. To show surjectivity, consider an arbitrary $Y\in \OPG(H).$ Define $\phi_{Y}:\mathcal{U}(\mathfrak{L}(\vec{a}) \to H$ by algebraically extending the mapping of generators $\phi_{Y}:x_{i}^{(n)} \mapsto y_i^{(n)}$. Then it is clear that $\phi_Y$ is an isomorphism and $\phi_Y(X) = Y \in \OPG(H)$.

    We now prove the second statement of the theorem. By \cite[Prop. 2.4]{F23},
    for any $X \in \OPG(H)$ we have $\mathcal{P}(H) \cong \mathfrak{L}(X)$.
    Then, because $\mathfrak{L}(X) = \CC X \oplus [\mathfrak{L}(X),\mathfrak{L}(X)]$
    \lucas{TODO: -define  $\mathfrak{L}(X)$?, add citation for this fact}we have
    \[
        \CC X \cong \mathcal{P}(H)/[\mathcal{P}(H),\mathcal{P}(H)].
    \]
    Fixing $X \in \OPG(H)$, we can then interpret an arbitrary choice of $Y \in \OPG(H)$ as some homogeneous change of basis of the quotient space $\mathcal{P}(H)/[\mathcal{P}(H),\mathcal{P}(H)]$ relative to $X$. Each homogeneous component $\vec{Y}^{(n)}$ then has the form
    \[
        \vec{Y}^{(n)} = M^{(n)}\vec{X}^{(n)} + \vec{B}^{(n)},
    \]
    for some $M^{(n)} \in \GL(a_n,\CC)$ and $\vec{B}^{(n)} = \left(b_1^{(n)},b_2^{(n)},\ldots,b_{a_n}^{(n)}\right)$, for some $b_i^{(n)} \in [\mathcal{P}(H),\mathcal{P}(H)]_n$. 
    
    Let us fix a homogeneous basis $\biguplus_{n \geq 1}\{v_1^{(n)},v_2^{(n)},\ldots,v_{p_n - a_n}^{(n)}\}$ of $[\mathcal{P}(H),\mathcal{P}(H)]$. Then $\vec{B}^{(n)}$ can be identified with an $a_n$ by $p_n-a_n$ matrix $K^{(n)}$, such that $b_i^{(n)} = \sum_{j = 1}^{p_n - a_n}K_{ij}^{(n)}v_j^{(n)}$. Then we define the $a_n$ by $p_n$ matrix
    \[
    A^{(n)} = \Bigl[M^{(n)} \big| K^{(n)}\Bigr]
    \]
    as desired. 
    
    Conversely, given a sequence $\left(A^{(n)}\right)_{n \geq 1} = \left([P^{(n)}|Q^{(n)}]\right)_{n \geq 1}$ with $A^{(n)} \in \textrm{Mat}_{a_n \times p_n}$ and $P^{(n)} \in \GL(a_n,\CC)$, we see that
    \[
    \vec{Z}^{(n)} = \biguplus_{n\geq 1} \left(z_1^{(n)},z_2^{(n)},\ldots,z_{a_n}^{(n)}\right) \text{ where } z_i^{(n)} = \sum_{j = 1}^{a_n}P_{ij}^{(n)}x_{j}^{(n)} + \sum_{k = 1}^{p_n - a_n}Q_{ik}^{(n)}v_k^{(n)}
    \]
    descends to a basis for $\mathcal{P}(H)/[\mathcal{P}(H),\mathcal{P}(H)]$.
    It follows from \cite[Corollary 2.6]{F23}\lucas{TODO:fill in number}
     that $Z = \biguplus_{n \geq 1} \vec{Z}^{(n)} \in \OPG(H)$,
     and it is clear the correspondence between sequences
     $\left(A^{(n)}\right)_{n \geq 1}$ and $\OPG(H)$ is one to one.
\end{proof}

%The above results prove the first part of Theorem \ref{thm:OPGmaps}. By definition of $\mathfrak{L}(\vec{a})$ we have that $A_n\in \GL(\CC)$ correspond to all isomorphism between two free Lie algebra $\mathfrak{L}(\vec{a})$ and $\mathfrak{L}(\vec{b})$ and by definition of $\vec{B}_n$ and $\mathcal{U}(-)$ we have that $\vec{Y}_n$ correspond to all isomorphism between two universal enveloping algebra. Therefore, the above proposition gives us part (i) of Theorem \ref{thm:OPGmaps}.

\begin{remark}
    Given two Hopf algebras $H$ and $K$ satisfying the criterion of Proposition~\ref{},\lucas{TODO:fill in number} we may present a method to compute an explicit isomorphism. Given graded sets $X$ and $Y$ generating $H$ and $K$ respectively, we apply Lemma~\ref{lemma:primitive generators} to obtain primitive generators $\mathbf{e}(X)$ and $\mathbf{e}(Y)$. Choosing bijections $f_n:\mathbf{e}(X)_n \to \mathbf{e}(Y)_n$ for all $n \geq 1$ we obtain a bijection of primitive generators $f:\mathbf{e}(X) \to \mathbf{e}(Y)$ which completely determines an isomorphism $\tilde{f}:H \to K$.
\end{remark}

\subsection{Examples that need a home}

\begin{example}\textit{A Cocommutative Hopf algebra on permutations.}
\lucas{TODO: cut this example down on detail, add references to the \FGCCHA papers where this appears: Mike says it may appear in two, Bergeron and possibly one other.  If there are good connections, we can keep it after Theorem~\ref{thm:a sequence}}

    Let $\vec{a}$ be the sequence determined by
    \[
    \sum_{n \geq 1}a_nt^n = 1 - \frac{1}{\sum_{m \geq 0}m!t^m}.
    \]
    Then $\vec{a} = (1,1,3,13,71,\ldots)$ corresponds to~\cite[A003319]{OEIS} which counts the number of indecomposable permutations of $[n]$. By \ref{thm: Hopf dimension isomorphic} there exists a \textsf{FGCCHA} of graded dimension $n!$, which we will construct explicitly.

    Let (placeholder) be the graded vector space spanned by permutations of $[n]$ in degree $n$. We will denote by $\pi = \pi_1\pi_2\cdots\pi_n$ the permutation of $[n]$ mapping $m \mapsto \pi_m$. For permutations $\sigma = \sigma_1\sigma_2\cdots\sigma_n$ of $[n]$ and $\tau = \tau_1\tau_2\cdots\tau_k$ of $[k]$, we define their concatenation product to be the permutation of $[n+k]$, 
    \[
    \sigma |\tau = \sigma_1\sigma_2\cdots\sigma_n(\tau_1 + n)(\tau_2+n)\cdots(\tau_k+n)
    \]
    which gives () the structure of a graded algebra. 
    
    The \textit{indecomposable permutations} are the permutations $\pi$ which cannot be written as the concatenation product of two non-trivial permutations, hence they are free generators for (). We now define the coproduct by declaring
    \[
    \Delta(\iota) = 1 \otimes \iota + \iota \otimes 1
    \]
    for all indecomposable permutations $\iota$, and extending algebraically. Under this coproduct () is endowed with the structure of a \textsf{FGCCHA}, with graded dimension $n!$.
\end{example}

\begin{example}
\lucas{TODO: This is our OPG Example~\ref{ex:OPGexample}, up to some re-writing and adaptation!}
    The Hopf algebra $\mathbf{CQSym}$ introduced in ~\cite{NT05} is a free graded connected cocommutative Hopf algebra, with a basis $\mathbf{P}^{\pi}$ indexed by non-crossing set partitions $\pi$, and is freely generated by $\{\mathbf{P}^{\pi}:\pi \text{ is atomic and non-crossing}\}$~\cite[Proposition 5.2]{NT05}.

    There is a sub-Hopf algebra $\mathbf{\Pi}^{\mathrm{NC}}$ of $\mathsf{NCSym}$ spanned by the power-sum basis elements $\mathbf{p}_{\pi}$ indexed also by non-crossing set partitions, and freely generated by $\{\mathbf{p}_{\pi}:\pi \text{ is atomic and non-crossing}\}$ which can be seen from the product and coproduct formula,
    \[
    \mathbf{p}_{\pi}\cdot \mathbf{p}_{\sigma} = \mathbf{p}_{\pi | \sigma}
    \]
    \[
    \Delta (\mathbf{p}_{\pi}) = \sum_{S \subseteq [\ell (\pi)]} \mathbf{p}_{\pi_S} \otimes \mathbf{p}_{\pi_{S^c}},
    \]
    where for $\pi = \{B_1,B_2, \ldots, B_{\ell(\pi)}\}$ and $S = \{s_1,s_2, \ldots, s_k\} \subseteq [\ell(\pi)]$, we define 
    \[
    \pi_{S} = \mathrm{st}\{B_{s_1},B_{s_2}, \ldots, B_{s_k}\}.
    \]
    
    For all $n \geq 0$, we have $\dim(\mathbf{CQSym}_n) = \dim(\mathbf{\Pi}^{\mathrm{NC}}_n) = c_n$ where $c_n$ is the $n$-th Catalan number. By Theorem~\ref{thm: Hopf dimension isomorphic} we have $\mathbf{CQSym} \cong \mathbf{\Pi}^{\mathrm{NC}}$, and remarkably due to the identity 
    \[
    \sum_{n \geq 0}c_nt^n = \frac{1}{1-\sum_{n \geq 1}c_{n-1}t^n},
    \]
    we see that the number of generators in degree $n$ for each algebra is $c_{n-1}$, or the number of atomic non-crossing partitions of $[n]$. 

    To realize an explicit isomorphism as outlined above, $\mathbf{e}(\mathbf{P}^{\pi})$ and $\mathbf{e}(\mathbf{p}_{\pi})$ running over all atomic non-crossing set partitions $\pi$ constitute free primitive generators for their respective algebras, and the map
    \[
    f:\mathbf{e}(\mathbf{P}^{\pi}) \mapsto \mathbf{e}(\mathbf{p}_{\pi})
    \]
    induces a Hopf algebra isomorphism $\tilde{f}:\mathbf{CQSym} \to \mathbf{\Pi}^{\mathrm{NC}}$
\end{example}

\subsection{Surjections}
\lucas{This section is bad and the statements are gross.  However, I have written enough to convince myself that it can be done, meaning we need to decide whether it should be done.}

This section attempts to prove the following statement, which is probably a general statement about free algebras.  It isn't the most interesting but does round out our picture.

\begin{thm}
Let $H$ and $K$ be \FGCCHAs and fix ordered primitive generating sets $X \in \OPG(H)$ and $Y \in \OPG(K)$.
\begin{enumerate}
\item There is a surjective homomorphism $\phi: H \to K$ only if $\veccard(X) \ge \veccard(Y)$.

\item For every surjective homomorphism $\phi: H \to K$, there exists elements $X' \in \OPG(H)$ and $Y' \in \OPG(K)$ for which
\[
\phi(x_{i}^{(n)}) = \begin{cases} y_{i}^{(n)} & \text{for $1 \le i \le |Y^{(n)}|$} \\ 0 & \text{otherwise} \end{cases}
\]

\end{enumerate}
\end{thm}
\begin{proof}
To prove the first statement, note that by~\ref{},
\[
\veccard(X) = \vecdim(\mathcal{P}(H) / [\mathcal{P}(H), \mathcal{P}(H)])
\qquad\text{and}\qquad
\veccard(Y) = \vecdim(\mathcal{P}(K) / [\mathcal{P}(K), \mathcal{P}(K)]).
\]
Now, any surjective homomorphism $\phi: H \to K$ induces a surjective linear map 
\[
(\mathcal{P}(H) / [\mathcal{P}(H), \mathcal{P}(H)] \to \mathcal{P}(K) / [\mathcal{P}(K), \mathcal{P}(K)],
\]
as $\phi$ must preserve primitives and the bracket.  By dimension consideration, we therefore have $\veccard(X) \ge \veccard(Y)$.

To prove the second, begin with our fixed element $X \in \OPG(H)$ and consider $\ker(\phi)$.  
If this kernel is generated by elements of $X$, then we have the desired property and are done.  
If not, take $n$ to be the smallest degree for which $\ker(\phi)$ is not generated by generators.  By assumption, $\ker(\phi)$ contains an element
\[
f = \sum_{S} a_{S} \prod_{(i, m) \in S} x_{i}^{(m)} \qquad \text{for $a_{S} \in \CC$},
\]
where the sum is over subsets $S$ of tuples $(i, m)$ with $1 \le i \le |X^{(m)}|$ for which $x_{i}^{(m)}$ is not contained in $\ker(\phi)$ and sum of the second item in each tuple of $S$ is $n$.
We need to show:
\begin{enumerate}
\item $f$ has at least one degree $n$ generator in it: otherwise, we can show $K$ isn't free by showing that it's dimension/generator sequence is wrong (using minimality assumption, lower degree parts of $\phi$ send generators to generators or zero, so we get a relation on generators of $K$)

\item Take $i$ to be minimal such that $x^{(n)}_{i} \in X^{(n)}$ appears with nonzero coefficient in some $f$ as above.  Then we can replace $x_{i}^{(n)}$ in our generating set by $f$ and still have a free generating set of $H$, but now $\phi$ sends this new generator to zero.  By induction, the number of degree $n$ generators which appear in nontrivial relations is reduced.  (note changing $x_{i}^{(n)} \to f$ can never introduce an $x_{j}^{(n)}$ for $j < i$ into a relation, since these don't appear in $f$).

\item Finally, it's possible our generating set isn't primitive anymore.  If we care about this (maybe we don't), we just hit the set with the Eulerian idempotent, which preserves Hopf algebra morphisms, and therefore the desired zeroing out property.

\end{enumerate}
\end{proof}

%\section{Classification of free cocommutative graded connected Hopf algebras} 
%
%Given a sequence $\vec{a} = (a_{1}, a_{2}, \ldots)$ of nonnegative integers, let $\mathfrak{L}(\vec{a})$ denote the free, graded Lie algebra with $a_{i}$ generators in degree $i$, and zero generators in degree zero.  The universal enveloping algebra
%\[
%\mathcal{U}\big(\mathfrak{L}(\vec{a})\big)
%\]
%is a cocommutative, graded, connected Hopf algebra, and we have shown that it is free, with Hilbert series
%\[
%\frac{1}{1 - a_{1}t - a_{2}t^{2} - a_{3}t^{3} - \cdots}.
%\]
%
%%\begin{thm}
%%Let $H=\bigoplus_{n\geq 0} H_n$ and $K=\bigoplus_{n\geq 0} K_n$ be two free graded connected cocommutative Hopf algebras. The following are equivalent.
%%\begin{enumerate}
%%\item $H\cong K$ as graded Hopf algebras. 
%%
%%\item $\dim(H_n)=\dim(K_n) \text{ for all } n\geq 0.$
%%\end{enumerate}   
%%\end{thm}
%%
%%\begin{proof}
%%Immediately we have (1)$\Rightarrow$(2). Assume that $\dim(H_n)=\dim(K_n) \text{ for all } n\geq 0,$
%%or equivalently $\mathcal{H}(H;t)=\mathcal{H}(K;t).$
%%Since both $H$ and $K$ are both freely generated, there exist graded sets
%%$X$ and $Y$ generating $H$ and $K$, respectively.
%%Define the series $\mathcal{G}(X;t)=\sum_{n\geq 0}|X_n|t^n$.
%%It follows from Proposition \ref{prop: free algebra hilbert series}
%%\mike{is this right?  I found "Prop. 3.2" in the text and
%%it no longer referred to the correct thing} that
%%\[
%%\mathcal{H}(H;t)=\frac{1}{1-\mathcal{G}(X;t)}\ \text{  and  }\ \mathcal{H}(K;t)=\frac{1}{1-\mathcal{G}(Y;t)}.
%%\]
%%Because $H$ and $K$ are connected, the degree zero terms of their Hilbert series $\dim(H_0)=\dim(K_0)=1$ are non-zero, hence the following equality of series
%%\[
%%\mathcal{G}(X;t)=1-\frac{1}{\mathcal{H}(H;t)}=1-\frac{1}{\mathcal{H}(K;t)}=\mathcal{G}(Y;t).
%%\]
%%We then have $|X_i|=|Y_i|$ for all $i\geq 0.$ Fix bijections $\varphi_i:X_i\to Y_i$ and extend
%%to a bijection $\varphi:X\to Y$. This graded bijection of generating sets induces a graded
%%Lie algebra isomorphism $\hat{\varphi}:\mathfrak{L}_X\to \mathfrak{L}_  Y$.\mike{How? reference?}
%%From Corollary~\ref{cor: free hopf primitives} we have
%%\[
%%\mathcal{P}(H)\cong \mathfrak{L}_X \text{ and } \mathcal{P}(K)\cong \mathfrak{L}_Y.
%%\]
%%We conclude by the Cartier-Milnor-Moore\mike{reference?} theorem that
%%\[
%%H\cong \mathcal{U}(\mathfrak{L}_X)\cong \mathcal{U}(\mathfrak{L}_Y)\cong K.
%%\]
%%\end{proof}
%
%
%\begin{thm}[Various Sources]
%Let $H$ be a free, cocommutative, graded, connected Hopf algebra, and let $\vec{a} = (a_{1}, a_{2}, \ldots, )$ count the number of free generators of $H$ in each degree.  
%\begin{enumerate}
%\item There is a graded isomorphism $H \cong \mathcal{U}\big(\mathfrak{L}(\vec{a})\big)$ of Hopf algebras.
%
%\item Every graded sub-Hopf algebra of $H$ is isomorphic to $\mathcal{U}\big(\mathfrak{L}(\vec{b})\big)$ for some sequence $\vec{b} = (b_{1}, b_{2}, \ldots)$.
%
%\item For every sequence $\vec{b} = (b_{1}, b_{2}, \ldots)$ with $b_{i} \le a_{i}$, $H$ has a graded sub-Hopf algebra isomorphic to $\mathcal{U}\big(\mathfrak{L}(\vec{b})\big)$.
%
%\end{enumerate}
%\end{thm}
%\begin{proof}
%Here is a sketch of what we need to formalize:
%\begin{enumerate}
%\item %Using Aliniaeifard--Thiem's result~\cite[Theorem 4.2]{AT22}, we only need to compare the Hilbert series of $H$ and $\mathcal{U}\big(\mathfrak{L}(\vec{a})\big)$ to prove this part.  \lucas{Should we write out their proof to convince ourselves it is true.  It is short.}
%
%Following Aliniaeifard--Thiem's result~\cite[Theorem 4.2]{AT22}, one only need to compare the Hilbert series of $H$ and $\mathcal{U}\big(\mathfrak{L}(\vec{a})\big)$.
%
%
%\item Let $G \subseteq H$ be a sub-Hopf algebra.  By (1), the Lie algebra of primitives $\mathcal{P}(H) \cong \mathcal{L}(\vec{a})$.  Thus $\mathcal{P}(G)$ is isomorphic to a subalgebra of $\mathcal{L}(\vec{a})$.  
%
%By~\cite[Theorem 2.2]{MSZ}, $\mathcal{P}(G)$ is a free Lie algebra, so there exists a sequence $\vec{b} = (b_{1}, b_{2}, \ldots)$ for which $\mathcal{P}(G) \cong \mathfrak{L}(\vec{b})$.
%%\lucas{Detail: show that this result plays well with the grading?}  
%By Milnor--Moore, $G \cong \mathcal{U}\big(\mathfrak{L}(\vec{b})\big)$.
%
%%Missing: what can we say about $\vec{b}$, i.e. the rank of $\mathcal{P}(G)$?  I really think it should satisfy $b_{i} \le a_{i}$, but free objects are weird so this needs a proof\lucas{Big detail!! This actually seems to be false}.  Since we are wrong here, we still could have a classification of sub-Hopf algebras by sequences $\vec{b}$, but we would need to figure out the correct constraints on $\vec{b}$.
%
%\item Here, we construct $\mathfrak{L}(\vec{b})$ as a subalgebra of $\mathfrak{L}(\vec{a})$ by taking $b_{i}$ of the free generators in degree $i$.
%\end{enumerate}
%\end{proof}
%
%Previously, we had thought that the condition in (3) above was sufficient, i.e. that every graded sub-Hopf algebra of $\mathcal{U}\big(\mathfrak{L}(\vec{a})\big)$ was isomorphic to $\mathcal{U}\big(\mathfrak{L}(\vec{b})\big)$ for some sequence $(b_{1}, b_{2}, \ldots)$ with $0 \le b_{i} \le a_{i}$ for all $i \ge 1$.  However, this is not true!  
%
%Consider the free, connected, cocommutative Hopf algebra which is freely generated as an algebra by four homogeneous elements of degree one, $x$, $y$, $z$, and $w$, so that
%\[
%H_{1} = \CC\operatorname{-span}\{x, y, z, w\}
%\]
%and
%\[
%H_{k} = \CC\operatorname{-span}\big\{ \text{products $m_{1} m_{2} \cdots m_{k}$} \;|\; m_{i} \in \{x, y, z, w\} \big\}
%\qquad\text{for $k \ge 2$}.
%\]
%By Proposition~\ref{prop: free algebra hilbert series}, $H$ has the Hilbert series
%\[
%\frac{1}{1 - 4t - 0 t^{2} - 0 t^{3} - \cdots }.
%\]
%
%The Lie algebra of primitives $\mathcal{P}(H)$ is the free Lie algebra $\mathfrak{L}_{\{x, y, z, w\}}$.  Now consider the Lie subalgebra
%\[
%M = \CC\operatorname{-span}\{[x, y]\} \subseteq \mathfrak{L}_{\{x, y, z, w\}},
%\]
%as well as its universal enveloping algebra $\mathcal{U}(M) \subseteq H$.
%By Theorem~\ref{}, this is a free Lie algebra generated by one element, and by dimension consideration we can take the free generator to be $[x, y]$.  
%Therefore by Theorem~\ref{}, the universal enveloping algebra $\mathcal{U}(M)$ is freely generated (as an algebra) by $[x, y]$.
%Now $[x, y] = xy - yx$ is a homogeneous element of degree $2$ in $H$, so as the subalgebra generated by $[x, y]$, the universal enveloping algebra $\mathcal{U}(M)$ is a graded sub-Hopf algebra and has Hilbert series
%\[
%\frac{1}{1 - 0t - t^{2} - 0t^{3} - \cdots}.
%\]
%Since $H$ has no free generators in degree $2$, this shows that sub-Hopf algebras may have more free generators in a given degree than the Hopf algebra which contains them.
%
%Is is also possible to obtain a sub-Hopf algebra which has more free generators (in total) than the Hopf algebra which contains it.  
%Consider the Lie subalgebra
%\[
%N = \langle [x, y], [x, z], [x, w], [y, z], [y, w], [z, w] \rangle \subseteq \mathfrak{L}_{\{x, y, z, w\}}.
%\]
%This is free on the set $\{[x, y], [x, z], [x, w], [y, z], [y, w], [z, w]\}$.  \lucas{Definitely true, but details needed.}  Repeating the argument we made for $\mathcal{U}(M)$, we have that $\mathcal{U}(N)$ is a graded sub-Hopf algebra of $H$ with six free generators in degree two, so that its Hilbert series is
%\[
%\frac{1}{1 - 0t - 6t^{2} - 0 t^{3} - \cdots }.
%\]
%
%\begin{question}
%Let $\vec{a} = (a_{1}, a_{2}, \ldots)$ be a sequence of nonnegative integers.  What characterizes the sequences $\vec{b} = (b_{1}, b_{2}, \ldots)$ for which $\mathcal{U}(\mathfrak{L}(\vec{a}))$ has a graded sub-Hopf algebra isomorphic to $\mathcal{U}(\mathfrak{L}(\vec{b}))$?
%\end{question}
%
%
%
%Given a sequence $\vec{a} = (a_1,a_2, \ldots)$ we have a formula for the dimension of the free Lie algebra with $a_i$ generators in each degree $i$~\cite[Theorem 2.2]{KK95}:
%\[
%\dim \mathfrak{L}(\vec{a})_n = \sum_{mk=n} \frac{1}{k}\mu (k) \mathbf{W}_m(\vec{a})
%\]
%where $\mu$ is the Mobius function, and setting $\lambda^!$ to be the multipllicity factorial and $\ell(\lambda)$ the length of $\lambda$,
%\[
%\mathbf{W}_m(\vec{a}) = \sum_{\lambda \vdash m} \frac{(\ell(\lambda)-1)!}{\lambda^!}a_{\lambda_1} \cdots a_{\lambda_{\ell}}.
%\]
%Define 
%\[
%P_n(\vec{a}) = \sum_{mk=n} \frac{1}{k}\mu (k) \mathbf{W}_m(\vec{a})
%\]
%so that $P_n(\vec{a}) = \dim \mathfrak{L}(\vec{a})_n$.
%
%\begin{theorem}
%\label{thm:subclassification1}
%    Let $\vec{a} = (a_{1}, a_{2}, \ldots)$ and $\vec{b} = (b_{1}, b_{2}, \ldots)$ be sequences of nonnegative integers. The following are equivalent.
%    \begin{enumerate}
%        \item $\mathcal{U}(\mathfrak{L}(\vec{a}))$ has a sub-Hopf algebra isomorphic to $\mathcal{U}(\mathfrak{L}(\vec{b}))$
%        \item $\mathfrak{L}(\vec{a})$ has a sub-Lie algebra isomorphic to $\mathfrak{L}(\vec{b})$.
%        \item $P_n(\vec{a}) \geq P_n(\vec{b})$ for all $n > 0$.
%    \end{enumerate}
%\end{theorem}
%
%\begin{proof}
%    The equivalence of (1) and (2) is clear. We will show (2) and (3) are equivalent. Suppose $\mathfrak{L}(\vec{a})$ has a sub-Lie algebra isomorphic to $\mathfrak{L}(\vec{b})$. Then
%    \[
%    P_n(\vec{b}) = \dim \mathfrak{L}(\vec{b})_n \leq \dim \mathfrak{L}(\vec{a})_n = P_n(\vec{a}).
%    \]
%
%    Now suppose $P_n(\vec{a}) \geq P_n(\vec{b})$ for all $n > 0$. We will construct a sub-Lie algebra $B \subseteq \mathfrak{L}(\vec{a})$ isomorphic to $\mathfrak{L}(\vec{b})$ by recursively choosing $b_n$-dimensional subspaces complementary to $[B,B]_n$ in each degree $n$. Let $V \subseteq \mathfrak{L}(\vec{a})$ be the subspace spanned by the free generators, so that $\mathfrak{L}(\vec{a}) = V \oplus [\mathfrak{L}(\vec{a}),\mathfrak{L}(\vec{a})]$. Note that we can recursively define 
%    \[ 
%    [B,B]_n = \sum_{k=1}^{n-1} [B_k,B_{n-k}].
%    \]
%
%    For $n=1$ we have $P_1(\vec{a}) = a_1$ and $P_1(\vec{b}) = b_1$, hence $a_1 \geq b_1$. Let $W_1$ be the span of $b_1$ of the degree 1 generators. Trivially $W_1$ is complementary to $[B,B]_1 = \{0\}.$
%    Suppose in each degree $1 \leq k < n$ we have chosen subspaces $W_k$ complementary to $[B,B]_k$. If $b_n \leq a_n$ we may simply choose a $b_n$-dimensional subspace of $V_n$. Otherwise, suppose $b_n > a_n$. Because $\mathfrak{L}(\vec{a})_n = V_n \oplus [\mathfrak{L}(\vec{a}),\mathfrak{L}(\vec{a})]_n$, we have
%    \[
%    \dim [\mathfrak{L}(\vec{a}),\mathfrak{L}(\vec{a})]_n = \dim \mathfrak{L}(\vec{a}) - \dim V_n = P_n(\vec{a}) - a_n.
%    \]
%    Because the spaces $B_k = W_k \oplus [B,B]_k$ are defined for $1 \leq k < n$, the space $[B,B]_n$ is defined recursively. Furthermore, $[B,B]_n$ is a subspace of $[\mathfrak{L}(\vec{a}),\mathfrak{L}(\vec{a})]_n$.
%    \\
%    \textit{Claim.} $\dim [B,B]_n = P_n(\vec{b}) - b_n$.
%    \\                    
%    
%    Given the claim, we have
%    \[
%    \dim \frac{[\mathfrak{L}(\vec{a}),\mathfrak{L}(\vec{a})]_n}{[B,B]_n} = P_n(\vec{a}) - a_n - \dim [B,B]_n \geq P_n(\vec{a}) - P_n(\vec{b}) + b_n - a_n \geq b_n - a_n. 
%    \]
%    Then we may choose $b_n - a_n$ linearly independent elements of $\frac{[\mathfrak{L}(\vec{a}),\mathfrak{L}(\vec{a})]_n}{[B,B]_n}$, and together with the $a_n$ generators in degree $n$ they span a $b_n$-dimensional subspace $W_n$ complementary to $[B,B]_n.$ Thus we have recursively defined 
%    \[
%    B = \bigoplus_{n\geq 1}W_n \oplus [B,B]_n = W \oplus [B,B].
%    \]
%    with $W = \bigoplus_{n\geq 1} W_n$ and $\dim W_n = b_n$.
%    
%    Clearly $B$ is closed under the Lie bracket, and is a sub-Lie algebra of $\mathfrak{L}(\vec{a}).$ By the Shirshov-Witt Theorem ~\cite[text]{S09}, $B$ is a free Lie algebra and is therefore generated by a basis for $W$, hence $B \cong \mathfrak{L}(\vec{b}).$
%\end{proof}
%
%\begin{proof}[Proof of Claim]
%Let $W_k$ be defined as above for $1 < k < n.$ Set $W_{<n} = \bigoplus_{k=1}^{n-1} W_k$ and let $\mathfrak{L}(W_{<n})$ be the free Lie algebra on $W_{<n}$  (this is equivalent to the free Lie algebra generated by a basis for $W_{<n}$). Since $\mathfrak{L}(W_{<n}) = W_{<n} \oplus [\mathfrak{L}(W_{<n}) , \mathfrak{L}(W_{<n})]$, we have $\mathfrak{L}(W_{<n})_1 = W_1 = B_1$, and recursively for $1 < k <n$ we have
%\[
%\mathfrak{L}(W_{<n})_k = W_k \oplus [\mathfrak{L}(W_{<n}) , \mathfrak{L}(W_{<n})]_k = W_k \oplus [B,B]_k = B_k
%\]
%therefore $[B,B]_n = [\mathfrak{L}(W_{<n}),\mathfrak{L}(W_{<n})]_n$.
%
%Define the sequence $\vec{b} \vert_n = (b_1,b_2, \ldots , b_{n-1}, 0, \ldots)$. Then $\dim \mathfrak{L}(W_{<n})_n = P_n(\vec{b}|_n)$. Because the only term of 
%\[ 
%P_n(\vec{b}) = \sum_{mk=n} \frac{1}{k}\mu (k) \sum_{\lambda \vdash m} \frac{(\ell(\lambda)-1)!}{\lambda^!}b_{\lambda_1} \cdots b_{\lambda_{\ell}}
%\]
%containing $b_n$ is with $m = n$ and $\lambda = (n)$ with a coefficient of $1$, we have $P_n(\vec{b}|_n) = P_n(\vec{b}) - b_n$. Therefore
%\[
%\dim [B,B]_n = \dim [\mathfrak{L}(W_{<n}),\mathfrak{L}(W_{<n})]_n = P_n(\vec{b}|_n) = P_n(\vec{b})-b_n.
%\]
%\end{proof}

%\section{A set of details to add}
%
%Aliniaeifard-Thiem \cite[Theorem 4.2]{AT22} give a theorem that says that two co-commutative Hopf algebras that
%have the same sequence of graded dimensions are isomorphic.  We should have a section here
%that says how to carry out this construction (outlined in the Cocalc summary for July 3).
%
%Let \(H\) and \(K\) be two free graded cocommutative Hopf algebras, such that \(\dim(H_n) = \dim(K_n)\) for all \(n \geq 0\).
%Aliniaeifard-Thiem  ~\cite[Theorem 4.2]{AT22}, guarantees that \(H \cong K\) and we outline here the method for constructing an 
%isomorphism \(f:H \to K\).  In general, the idea is to find the free primitive generators of \(H\) and \(K\), and map the primitive generators of 
%each graded component \(H_n\) to the primitive generators of the  coresponding graded component \(K_n\).
%
%Aguiar-Lauve \cite{AL15} give the Eulerian idempotent operator
%\[
%    \mathbf{e} = \sum_{k=1}^\infty \frac{(-1)^{k-1}}{k} m^{(k-1)}\circ \Delta^{(k-1)}
%\]
%which for any cocommutative Hopf algebra projects onto the space of primitive elements. Since \(H\) and \(K\) are cocommutative
%we can use \(\mathbf{e}\) to compute  \(\mathcal{P}(H)\) and \(\mathcal{P}(K)\).
%Then, by ~\ref{lemma:primitive generators}, there exists \(X \subseteq \mathcal{P}(H)\) such that \(H\) is generated by \(X\)
%and  \(Y \subseteq \mathcal{P}(K)\) such that  \(K\) is  generated by \(Y\).
%
%In order to distinguish the free primitive generators \(X\)
%from the primitive elements \(\mathcal{P}(H)\) consider the decomposition \[
%    \mathcal{P}(H) = V \oplus [\mathcal{P}(H), \mathcal{P}(H)]
%\] where \([\mathcal{P}(H), \mathcal{P}(H)] = \CC\operatorname{-span}\{[p1,p2] \: \vert \: p_i \in \mathcal{P}(H)\}\) and
%\(V\) is a graded subspace of \(\mathcal{P}(H)\). Since \(\mathcal{P}(H)\) forms a Lie algebra under the commutator bracket,
%then, by ~\cite[Prop 2.2]{F23}, \(\mathcal{P}(H)\) is generated by a basis of \(V\). Hence \(X\) is a basis of \(V\). Therefore, in general,
%to find the set of primitive generators of \(H\), first choose a graded subspace \(V \subseteq \mathcal{P}(H)\), then choose a basis for \(V\).
%Alternatively, letting \(V = \frac{\mathcal{P}(H)}{[\mathcal{P}(H), \mathcal{P}(H)]}\),
% any subset \(X' \subseteq \mathcal{P}(H)\) that descends to a basis of \(V\), will be a set primitive generators of \(H\).
%
%Let \(p_i\) be a basis of \(V\) and let  \(A \in GL_{|I|}(\CC)\) be an invertible linear transformation.
%Then, for some  \(b_i \in [\mathcal{P}(H), \mathcal{P}(H)]\),
%we can characterize all possible bases for \(V\) as \[
%    q_i = \sum_jA_{ij}p_i + b_i.
%\]
%Hence we find the free primitive generators \(X\) of \(H\) and \(Y\) of \(K\), and, as in the proof of
%\cite{AT22}, fix bijections \(f_i: X_i  \to Y_i\) and extend to a bijection
%\(f:X \to Y\).

%\section{Empty this section out}
%\lucas{TODO: This section is almost empty!  Let's finish clearing it out and delete it.}

%\mike{It would be nice to have an example of a sequence $\vec{h}$ that
%is not obvious that it doesn't have a Hopf algebra.
%For instance, $\vec{h} = (1,2,2,2,2,\ldots)$ is there a Hopf algebra?
%If yes, find me a different example.  If no, apply this theorem to explain why.}

\section{Classifying SubHopf algebras by primitive dimensions}

Let $H$ be a free graded connected cocommutative Hopf algebra.  
In this section we construct all the isomorphism classes of graded Hopf algebra which occur as sub-Hopf algebras of $H$.  
To begin, recall the function $\phi_{p, a}$ on $\CC^{\NN}$ defined by
\[
\phi_{p, a}(\vec{p}) = \big( \sum_{\lambda \vdash n} (-1)^{\ell(\lambda) - 1} \prod_{d \ge 0} \binom{p_{d}}{m_{d}(\lambda)} \big)_{n = 1}^{\infty}
\]
and recall that for any nonnegative sequence $\vec{a} = (a_{1}, a_{2}, \ldots)$, we write $\mathfrak{L}(\vec{a})$ for the free Lie algebra generated by $a_{n}$ elements in degree $n$.

\begin{thm}
\label{thm:subclassification2}
Let $H$ be a free graded connected cocommutative Hopf algebra and let $\vec{p} = \vecdim(\mathcal{P}(H)_{n})$.  Then there is a bijection
\[
\begin{array}{rcl}
\left\{\begin{array}{c}
\text{Isomorphism classes of} \\
\text{sub-Hopf algebras $G \subseteq H$}
\end{array}\right\}
& \leftrightarrow & 
\left\{\begin{array}{c}
\text{Sequences $\vec{q} = (q_{1}, q_{2}, \ldots)$ such } \\
\text{that $\vec{p} \ge \vec{q} \ge \vec{0}$ and $\phi_{p, a}(\vec{q}) \ge \vec{0}$} 
%\\\sum_{\lambda \vdash n} (-1)^{\ell(\lambda) - 1} \prod_{d \ge 0} \binom{q_{d}}{m_{d}(\lambda)} \ge 0
\end{array}\right\} \\[2em]
G & \mapsto & \vecdim\big(\mathcal{P}(G)\big) \\
\mathcal{U}(\mathfrak{L}(\phi_{p, a}(\vec{q}))) & \mapsfrom & \vec{q}
\end{array}
\]
\end{thm}

The proof of Theorem~\ref{thm:subclassification2} will construct a sub-Hopf algebra in $H$ isomorphic to each $\mathcal{U}(\mathfrak{L}(\phi_{p, a}(\vec{q})))$.  
Before completing the proof we will give several motivating examples and intermediate results.

\begin{example}
We illustrate a counterintuitive aspect of containment between \FGCCHAs. Let $H$ and $K$ be \FGCCHAs generated by graded sets with respective graded cardinalities $\vec{a}=(0,1,0,0,\dots)$ and $\vec{b}=(1,0,0,0,\dots)$, and further denote $\vec{h} = \vecdim(H)$ and $\vec{k} = \vecdim(K)$.
Using Proposition~\ref{prop:sequences}, we can compute that
\[
h_n = \begin{cases} 1 & \text{if $n$ is even,} \\ 0 & \text{if $n$ is odd,} \end{cases}
\]
as $a_{\beta} \neq 0$ implies that $\beta = (2, \ldots, 2)$.  Similarly, we can compute that
\[
k_{n} = 1 \qquad\text{for all $n \ge 1$}.
\]
Therefore, $\vec{h} \le \vec{k}$, so one might expect that $K$ should contain a subHopf algebra isomorphic to $H$.  However, Theorem~\ref{thm:subclassification2} tells us this is not the case, as $\vec{a}$ and $\vec{b}$ are incomparable.
\end{example}

\begin{example}
\label{ex:NSymInCxy}
We apply the theorem to our running example: let $H = \CC\langle x, y \rangle$,
so that $\vec{p} = \vecdim(\mathcal{P}(H))$ is the sequence
$\vec{p} = (2, 1, 2, 3, 6, 9, 18, 30,\ldots)$~\cite[A001037]{OEIS} given by 
\[
p_{n} = \frac{1}{n} \sum_{d | n} \mu(n/d) 2^{d}.
\]

Then we can take 
\[
\vec{q} = \vec{p} - (1, 0, 0, \ldots)
\]
so that clearly $\vec{p} \ge \vec{q} \ge \vec{0}$.  
The sequence $\vec{q}$ is identified in~\cite[A059966]{OEIS}, and direct computation then gives that
\[
\phi_{p, a}(\vec{q}) = (1, 1, 1, \ldots) \ge \vec{0},
\]
so this sequence determines a subalgebra of $\mathcal{U}(\mathfrak{L}(1, 1, 1, \ldots))$
that is freely generated by one primitive element in each degree. 
We have already\lucas{TODO: reference earlier NSym Example} identified we have identified $\mathcal{U}(\mathfrak{L}(1, 1, 1, \ldots)) \cong \mathsf{NSym}$, we conclude that $H$ has a Hopf subalgebra isomorphic to $\mathsf{NSym}$.
\end{example}

We now state and prove an intermediate result in the proof of Theorem~\ref{thm:subclassification2}; the proof of the theorem follows.  Given a graded Lie algebra $L$, the (graded) derived subalgebra of $L$ is 
\[
[L, L] = \bigoplus_{n \ge 1} [L, L]_{n}
\qquad\text{where}\qquad
[L, L]_{n} = \CC\operatorname{-span}\{\text{degree-$n$ commutators of $L$}\}.
\]

\begin{lemma}
\label{lem:deriveddimension}
Let $\vec{a} = (a_{1}, a_{2}, \ldots)$ be a sequence of nonnegative integers.  Then
\[
\vecdim([\mathfrak{L}(\vec{a}), \mathfrak{L}(\vec{a})]) = \phi_{a, p}(\vec{a}) - \vec{a}.
\]
\end{lemma}
\begin{proof}
Using the fact that $\phi_{a, p}(\vec{a}) = \vecdim(\mathfrak{L}(\vec{a}))$, it is sufficient to show that
\[
\vec{a} = \vecdim\left( \mathfrak{L}(\vec{a}) \big/ [\mathfrak{L}(\vec{a}), \mathfrak{L}(\vec{a})] \right).
\]
By definition, $\vec{a}$ enumerates a fixed choice $X$ of free homogeneous generators for $\mathfrak{L}(\vec{a})$.  
Since free generators have no relations with lower-degree terms, the elements $X$ are linearly independent modulo the derived sub-Lie algebra $[\mathfrak{L}(\vec{a}), \mathfrak{L}(\vec{a})]$, so the dimension of the degree $n$ part of the quotient must be at least $a_{n}$.  
However, any element of homogeneous degree $n$ in $\mathfrak{L}(\vec{a})$ is,
by definition, a linear combination of some element of
$[\mathfrak{L}(\vec{a}), \mathfrak{L}(\vec{a})]$ and the degree $n$
generators, so the dimension of the quotient is at most $a_{n}$.
\mike{TODO: Mike can we delete this? two assertions are being made in this paragraph that
I think need to be justified: 1. the elements are lin ind mod $[\mathfrak{L}(\vec{a}), \mathfrak{L}(\vec{a})]$
and 2. $r \in \mathfrak{L}(\vec{a})$ implies $r = s + t$ with 
$s \in [\mathfrak{L}(\vec{a}), \mathfrak{L}(\vec{a})]$ and $t$ primitive}
\end{proof}


\begin{proof}[Proof of Theorem~\ref{thm:subclassification2}]
We first show that the given map is injective.  Suppose that $J$ and $K$ are sub-Hopf algebras of $H$ with the property that $\vecdim(\mathcal{P}(J)) = \vecdim(\mathcal{P}(K))$.  By Proposition~\ref{prop:SequencesAsHopfStructures}, we have
\[
\vecdim(J) = \phi_{p, h}\big(\vecdim(\mathcal{P}(J))\big)
\qquad\text{and}\qquad
\vecdim(K) = \phi_{p, h}\big(\vecdim(\mathcal{P}(K))\big),
\]
so that $\vecdim(J) = \vecdim(K)$.  We therefore conclude by Theorem~\ref{thm:a sequence} that $J \cong K$.
\mike{TODO: is this referring to the right theorem?}

Now, suppose that $\vec{q}$ is a sequence satisfying the conditions that $\vec{p} \ge \vec{q} \ge \vec{0}$ and $\vec{b} = \phi_{p, a}(\vec{q}) \ge \vec{0}$.  
We will construct a sub-Hopf algebra $K$ of $H$ which is isomorphic to $\mathcal{U}(\mathfrak{L}(\vec{b}))$; by Proposition~\ref{prop:sequences} this will show that the given map is surjective, completing the proof.  

In order to do so, we construct a tower of sub-Lie algebras 
\[
0 = L^{(1)} \subseteq L^{(2)} \subseteq L^{(3)} \subseteq \cdots \subseteq \mathcal{P}(H)
\]
such that 
\[
L^{(n)} \cong \mathfrak{L}(b_{1}, \ldots, b_{n-1}, 0, 0, \ldots).
\]
Taking $L^{(\infty)}$ to be the union of the $L^{(n)}$, we obtain $K$ as $\mathcal{U}(L^{(\infty)}) \subseteq \mathcal{U}(\mathcal{P}(H)) = H$.

Our construction is inductive with base case $L^{(1)} = 0$.  
Begin by assuming that $L^{(n)}$ has been constructed.  
By Lemma~\ref{lem:deriveddimension} and our inductive hypothesis,
\[
\vecdim([L^{(n)}, L^{(n)}])
 = 
\phi_{a, p}(b_{1}, \ldots, b_{n-1}, 0, 0, \ldots) - (b_{1}, \ldots, b_{n-1}, 0, 0, \ldots).
\]
Considering the $n$th term, we see that $\vecdim(L^{(n)})_{n} = \vecdim([L^{(n)}, L^{(n)}])_{n}$, and we can further deduce---after a careful examination of the definition of $\phi_{a, p}$---that

\begin{equation}
\label{eq:inductivedimensionequality}
\phi_{a, p}(b_{1}, \ldots, b_{n-1}, 0, 0, \ldots)_{n}
=
\phi_{a, p}(\vec{b}) - b_{n}
=
q_{n} - b_{n}.
\end{equation}
Therefore,
\[
\vecdim\left( \mathcal{P} \big/ [L^{(n)}, L^{(n)}] \right)_{n}
= p_{n} - (q_{n} - b_{n})
= b_{n} + (p_{n} - q_{n})
\ge b_{n}.
\]

The preceding inequality shows that we can choose $b_{n}$ linearly independent elements $\{x^{(n)}_{1}, \ldots, x^{(n)}_{b_{n}}\}$ of $\mathcal{P}(H)_{n}$ which remain linearly independent modulo $[L^{(n)}, L^{(n)}]$ and define
\[
L^{(n+1)} = \langle L^{(n)}, x^{(n)}_{1}, \ldots, x^{(n)}_{b_{n}} \rangle.
\]
Since $L^{(n+1)}$ is a sub-Lie algebra of a free Lie algebra, namely $\mathcal{P}(H)$,~\cite[Theorem 2.2]{MSZ} states that $L^{(n+1)}$ is also free.  
Moreover, as $L^{(n+1)}$ is generated by the generators of $L^{(n)}$ and homogeneous elements of degree $n$, we deduce that
\[
L^{(n+1)} \cong \mathfrak{L}(b_{1}, \ldots, b_{n-1}, b_{n}', 0, \ldots )
\qquad\text{for some $b_{n}' \le b_{n}$}.
\]
Moreover, using Lemma~\ref{lem:deriveddimension} and the definition of $L^{(n+1)}$,
\begin{align*}
b_{n}' &= \dim\left( L^{(n+1)}_{n} \big/ [L^{(n+1)}, L^{(n+1)}]_{n}\right) \\
&= \dim\left(  \CC\operatorname{-span}\{x^{(n)}_{1}, \ldots, x^{(n)}_{b_{n}}\} \oplus [L^{(n)}, L^{(n)}]_{n} \big/ [L^{(n)}, L^{(n)}]_{n} \right) \\
&= b_{n}. \qedhere
\end{align*}
\end{proof}

%\begin{example}
%Building on Example~\ref{ex:NSymInCxy}, we construct an isomorphic copy of $\mathsf{NSym}$ inside of $H = \CC\langle x, y \rangle$ in the manner of the above proof.  
%This means that for each degree $n \ge 1$, we will choose a primitive element $x^{(n)}_{1}$ in degree $n$ that is linearly independent of all commutators in generators $x_{1}^{(i)}$ for $i < n$, and then define
%\[
%L = \langle x_{1}^{(1)}, x_{1}^{(2)}, x_{1}^{(3)}, \ldots \rangle. 
%\]
%
%In degree $1$, there are no constrains and we choose 
%\[
%x_{1}^{(1)} = x.
%\]
%In higher degrees, we let
%\[
%x_{1}^{(n)} = \underbrace{[y, [y, \cdots [y}_{\text{$n-1$ times}}, x] \cdots]].
%\]
%To see that $x_{1}^{(n)}$ is linearly independent of all commutators in $x_{1}^{(i)}$ for $i < n$, we consider the number of times $x$ occurs in any generator or bracket of generators.  
%This will be $1$ if we have a generator and $k$ for a bracket of $k$ generators, and moreover remains constant on terms in any Jacobi relation or anti-commutation relation.  
%Thus, $L_{n}$ is graded by this statistic, with the degree $1$ component being the span of $x_{1}^{(n)}$, and the sum of components of degree $2$ or greater comprising all commutators of the $x_{1}^{(i)}$ for $i < n$.  
%\end{example}
%
%\begin{example}
%\lucas{GROUP TODO: Examples from earlier versions....do we keep?}
%\felix{If we want to talk about the idea we had in first place about the sequences then this can be a good counterexample otherwise I think we can let it go}
%\lucas{I generally agree, though this was a fun example to work out.}
%The Lie algebra of primitives $\mathcal{P}(H)$ is the free Lie algebra $\mathfrak{L}_{\{x, y, z, w\}}$.  Now consider the Lie subalgebra
%\[
%M = \CC\operatorname{-span}\{[x, y]\} \subseteq \mathfrak{L}_{\{x, y, z, w\}},
%\]
%as well as its universal enveloping algebra $\mathcal{U}(M) \subseteq H$.
%By Theorem~\ref{}, this is a free Lie algebra generated by one element, and by dimension consideration we can take the free generator to be $[x, y]$.  
%Therefore by Theorem~\ref{}, the universal enveloping algebra $\mathcal{U}(M)$ is freely generated (as an algebra) by $[x, y]$.
%Now $[x, y] = xy - yx$ is a homogeneous element of degree $2$ in $H$, so as the subalgebra generated by $[x, y]$, the universal enveloping algebra $\mathcal{U}(M)$ is a graded sub-Hopf algebra and has Hilbert series
%\[
%\frac{1}{1 - 0t - t^{2} - 0t^{3} - \cdots}.
%\]
%Since $H$ has no free generators in degree $2$, this shows that sub-Hopf algebras may have more free generators in a given degree than the Hopf algebra which contains them.
%
%Is is also possible to obtain a sub-Hopf algebra which has more free generators (in total) than the Hopf algebra which contains it.  
%Consider the Lie subalgebra
%\[
%N = \langle [x, y], [x, z], [x, w], [y, z], [y, w], [z, w] \rangle \subseteq \mathfrak{L}_{\{x, y, z, w\}}.
%\]
%This is free on the set $\{[x, y], [x, z], [x, w], [y, z], [y, w], [z, w]\}$.  \lucas{Definitely true, but details needed.}  Repeating the argument we made for $\mathcal{U}(M)$, we have that $\mathcal{U}(N)$ is a graded sub-Hopf algebra of $H$ with six free generators in degree two, so that its Hilbert series is
%\[
%\frac{1}{1 - 0t - 6t^{2} - 0 t^{3} - \cdots }.
%\]
%\end{example}
%

%\begin{proposition}
%    Let $H$ be a free graded connected cocommutative Hopf algebra, let $\mathbf{g} = (g_1,g_2,\ldots )$ be a sequence of nonnegative integers. There exists a sub-Hopf algebra $G \leq H$ with $\dim G_n = g_n$ for all $n \geq 0$ if and only if the following hold:
%    \begin{itemize}
%        \item[i)] $\displaystyle \sum_{\beta \vDash n} (-1)^{\ell(\beta)-1} g_\beta \geq 0 \text{ for all } n \geq 0.$
%        \item[ii)] $\displaystyle \sum_{d|n} \sum_{\beta \vdash d} \frac{d\cdot \mu(n/d) (-1)^{\ell(\beta)-1}}{n \cdot \ell(\beta)} g_\beta \leq \sum_{d|n} \sum_{\beta \vdash d} \frac{d\cdot \mu(n/d) (-1)^{\ell(\beta)-1}}{n \cdot \ell(\beta)} h_\beta$ for all $n \geq 0$
%    \end{itemize}
%\end{proposition}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\lucas{TODO: We need a better bibliography!}
\bibliographystyle{plain}
\bibliography{bibliography}{}


 \begin{table}[h]
    \hspace*{-1cm} 
    \begin{tabular}{|l|l|l|l||l|l|l||l|l|l|l||l|l|l|}
        \hline
        Sequence & h sequence & a sequence & p sequence \\
        \hline
        \href{https://oeis.org/A000108}{Catalan} & \(1, 2, 5, 14\) & 
        \(1, 1, 2, 5, 14, 42, 132, 429, 1430\) & 
        \(1, 1, 3, 8, 25, 75, 245, 800, 2700\) \\
        
        %\hline
        %\href{https://oeis.org/A001764}%{Ternary Trees} & \(1, 3, 12, 55,273\) & 
        %\(1, 2, 7, 30, 143, 728, 3876, %21318, 120175, 690690\) & 
        %\(1, 2, 9, 40, 200, 1026, 5537, %30624, 173583, 1001400\)
        %\\
        
        \hline
        \href{https://oeis.org/A001006}{Motzkin}  & \(1, 2, 4, 9, 21\) & 
        \(1, 1, 1, 2, 4, 9, 21, 51, 127, 323\) & 
        \(1, 1, 2, 4, 10, 22, 56, 136, 348, 890\)

          \\
        \hline
        \href{https://oeis.org/A000045}{ Fibonacci}
        & \(1, 1, 2, 3,5\) & 
        \(1, 0, 1, 0, 1, 0, 1, 0, 1, 0\) & 
        \(1, 0, 1, 1, 2, 2, 4, 5, 8, 11\)

         \\
        \hline
        \href{https://oeis.org/A000073}{Tribonacci} & \(0, 1, 1, 2, 4\) & 
        \(0, 1, 1, 1, 2, 3, 4, 6, 9, 13\) & 
        \(0, 1, 1, 1, 3, 4, 8, 13, 23, 38\)

             \\
        \hline
        \href{https://oeis.org/A000032}{Lucas} & \(2, 1, 3, 4, 7\) & 
        \(2, -3, 7, -13, 27, -53, 107, -213, 427, -853\) & 
        \(2, -2, 3, -2, 8, -9, 22, -25, 64, -94\)

         \\
        \hline
        \href{https://oeis.org/A000110}{Bell}
         & \(1, 2, 5, 15, 52\) & 
        \(1, 1, 2, 6, 22, 92, 426, 2146, 11624, 67146\) & 
        \(1, 1, 3, 9, 34, 135, 610, 2965, 15612, 87871\)


   \\
      
        \hline
    \end{tabular}
    \caption{Sequences}
    \label{tab:sequences}
\end{table}

Rooted trees (Grossman-Larson, Connes-Kreimer),
ordered trees (Grossman-Larson),
heap-ordered trees (Grossman-Larson),
planar binary trees (Loady-Ronco), permutations (Malvenuto-Reutenauer, Aguiar-Sottile),
compositions (QSym/NSym), odd compositions (Peak algebra of Stembridge),
parking functions ($(n+1)^{n-1}$).
Set partitions.

\end{document}
%